### 2020-03-25 raport.Rmd
Pierwszy ogląd o co chodzi
Z Messengera:
Mam złe wieści
Walczyłem teraz jakoś z godzinę-półtorej próbując bardziej ogarnąć o co w tym wszystkim chodzi i jest tu trochę wyższej matmy którą słabo rozumiem [albo mam mocne braki z algebry, zapewne pół na pół]

Raport.Rmd i raport.html to moje dzieło - proponuję po prostu prezentować .Rmd, ew. z wstawionymi obrazkami

źródła.txt to linki które się przydadzą - trochę info, jakaś dokumentacja i... podsumowanie autora pakietu, które także pobrałem do .Rmd w podsumowanieTworcy.Rmd i wywołałem

---
title: "softImpute"
author: "Jakub Kosterna, Dawid Przybyliński, Hanna Zdulska"
date: "25 marca 2020"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Wprowadzenie

## 1.1. Geneza

Ten genialny pakiet to doskonałe narzędzie do uzupełnienia macierzy liczbami przy użyciu **nuclear norm regularization**, znanej także jako **trace norm regularization** (pl: *regularyzacja norm śladowych*). Biblioteka implementuje funkcję *softImpute()* z dwoma przydatnymi algorytmami:

1. Iteracyjne obliczenie soft-thresholded SVD z wypełnionej macierzy (opcja *type="svd"*)

2. Uzupełnianie braków ramki szacunkami korzystając z alternating ridge regression (opcja *type="als"*)

Zaletą pakietu jest fakt, że może obsługiwać zarówno małe, jak i bardzo duże macierze.

## 1.2. Norma śladowa - przypomnienie

No dobra, ale czym właściwie była ta *trace norm*?

Jest to suma wartości bezwzględnych wartości własnych macierzy,

$||A||_{1} = trace((\sqrt{(X^{*}X)}))$

# 2. Działanie

## 2.1. Odczyt

```{r odczyt, message = FALSE, warning = FALSE}
# install.packages("softImpute") # if not installed
library(softImpute)
```

## 2.2. Przykład

Wygenerujmy losową macierz 5 x 6 i usuńmy losowe 7 wartości.

```{r przyklad1_1}
set.seed(123)

X = matrix(rnorm(30), 5, 6)
X
X_del = X
X_del[sample(1:30, 7, replace = FALSE)] = NA
X_del
```

Zobaczmy wynik działania *soft_impute()*. W tym celu utworzymy zmienną *fits* generującą dopasowanie w zależności od macierzy z usuniętymi wartościami, a następnie wypełnimy ją z pomocą funkcji *complete()*.

```{r przyklad1_2}
fits = softImpute(X_del, trace = TRUE, type="svd")
complete(X_del, fits)
```

Jak widać funkcja poradziła sobie całkiem dobrze w przypadku losowo wygenerowanych wartości z rozkładu normalnego.

### 2020-03-30 raport.Rmd
introduction

@@
   html_document:
+    df_print: paged
     toc: true
-    toc_depth: 2
+    toc_float: true
+    number_sections: true
 ---
@@
 
-# 1. Wprowadzenie
+# Wprowadzenie
 
-## 1.1. Geneza
+## Geneza i zawartość
 
-Ten genialny pakiet to doskonałe narzędzie do uzupełnienia macierzy liczbami przy użyciu **nuclear norm regularization**, znanej także jako **trace norm regularization** (pl: *regularyzacja norm śladowych*).
-Biblioteka implementuje funkcję *softImpute()* z dwoma przydatnymi algorytmami:
+SoftImpute to nieduży i przydatny pakiet do uzupełniania macierzy, stworzoną przez Trevor Hastiego(Stanford, autor wielu książek o statystce).
+Wykorzystuje do tego **nuclear norm regularization**, znanej także jako **trace norm regularization** (pl: *regularyzacja norm śladowych*).
+Biblioteka implementuje:
+
+  * funkcję *softImpute()* z dwoma przydatnymi algorytmami:
 
@@
 
-2. Uzupełnianie braków ramki szacunkami korzystając z alternating ridge regression (opcja *type="als"*)
+    2. Uzupełnianie braków ramki szacunkami korzystając z alternating ridge regression(zwane popularniej w Europie Tikhonov regularization) (opcja *type="als"*)
 
-Zaletą pakietu jest fakt, że może obsługiwać zarówno małe, jak i bardzo duże macierze.
+  * biscale() - zoptymalizowane scale() działające również na niekompletnej macierzy, działające też szybko na bardzo dużych macierzach
 
-## 1.2. Norma śladowa - przypomnienie
+  * dwie nowe klasy dla macierzy: SparseplusLowRank i Incomplete
+
+## Zalety i wady
+
+Pakiet z założenia służy do uzupełniania macierzy, wymaga on nas konwersji z data.frame na klasę matrix. Operuje tylko na macierzach, co jest równocześnie zaletą - potrzebuje tylko pakiet Matrix:
+```{r}
+require(deepdep)
+plot_dependencies(deepdep("Amelia", depth = 2))
+plot_dependencies(deepdep("missMDA", depth = 2))
+plot_dependencies(deepdep("missForest", depth = 2))
+plot_dependencies(deepdep("VIM", depth = 2))
+plot_dependencies(deepdep("mice", depth = 2))
+plot_dependencies(deepdep("softImpute", depth = 2))
+```
+
+Zaletą pakietu jest fakt, że może obsługiwać zarówno małe, jak i bardzo duże macierze w relatywnie krótkim czasie, dzięki użyciu klasy Incomplete.
+Dla przykładu `softImpute` bez problemu dopasuje macierz rzędu 100 do danych Netflixa (480,189 użytkowników x 17,770 filmów co daje około 8.6 * 10^9 obserwacji, przy czym średnio jeden użytkownik ocenia tylko 200 filmów, co daje tylko 1.2% obserwacji ~10^8) w 3.3h
+
+# Teoria
 
@@
 
-# 2. Działanie
+# Działanie
 
-## 2.1. Odczyt
+## Odczyt
 
@@
 
-## 2.2. Przykład
+## Przykład
 
@@
 Jak widać funkcja poradziła sobie całkiem dobrze w przypadku losowo wygenerowanych wartości z rozkładu normalnego.
+
+# Bibliografia - poczytaj wiecej
+ * [Mazumder et al (2010)](http://web.stanford.edu/~hastie/Papers/mazumder10a.pdf)
+ * [Hastie et al (2014)](http://arxiv.org/abs/1410.2596)

### 2020-03-31 raport.Rmd
algebra

@@
 ```{r}
-require(deepdep)
-plot_dependencies(deepdep("Amelia", depth = 2))
-plot_dependencies(deepdep("missMDA", depth = 2))
-plot_dependencies(deepdep("missForest", depth = 2))
-plot_dependencies(deepdep("VIM", depth = 2))
-plot_dependencies(deepdep("mice", depth = 2))
-plot_dependencies(deepdep("softImpute", depth = 2))
+#require(deepdep)
+#plot_dependencies(deepdep("Amelia", depth = 2))
+#plot_dependencies(deepdep("missMDA", depth = 2))
+#plot_dependencies(deepdep("missForest", depth = 2))
+#plot_dependencies(deepdep("VIM", depth = 2))
+#plot_dependencies(deepdep("mice", depth = 2))
+#plot_dependencies(deepdep("softImpute", depth = 2))
 ```
@@
 
-# Teoria
+## Jak działa algorytm? 
+Niech $X$  będzie dużą $m\times n$ macierzą z wiloma brakującymi wartościami.
+Niech $\Omega$ zawiera pary indeksów $(i,j)$ gdzie $X$ był zaobserwowany, i niech $P_\Omega(X)$ oznacza macierz tylko z  punktami z $\Omega$, a w pozostałych miejscach ma wartości 0.
+Zatem jeśli $X$ ma brakujące wartości w $\Omega^\perp$, $P_\Omega(X)$ by oznaczyło NA jako 0.
 
-No dobra, ale czym właściwie była ta *trace norm*?
+Rozważmy wyrażenie
+ $$\min_M\frac12\|P_\Omega(X)-P_\Omega(M)\|^2_F+\lambda\|M\|_*,$$
+ gdzie $\|M\|_*$ to suma wartości singularnych macierzy $M$ (nucelar norm, trace norm).
 
-Jest to suma wartości bezwzględnych wartości własnych macierzy,
+ Jeśli $\widehat M$ jest rozwiązaniem plroblemu, to musi spełniać następujący warunek:
+ $$ {\widehat M}=S_\lambda(Z)$$ 
+ Gdzie $$Z=P_\Omega(X)+P_{\Omega^\perp}(\widehat M).$$
+ Operator $S_\lambda(Z)$ dla macierzy $Z$ wykonuje następujące instrukcje:
 
-$||A||_{1} = trace((\sqrt{(X^{*}X)}))$
+ 1. Znajduje rozkład SVD dla $Z=UDV^T$, oraz $d_i$ - wartości singularne $D$.
+ 2. Zaszumia wartości singularne: $d_i^*= (d_i-\lambda)_+$.
+ 3. Tworzy nową macierz $Z$: $S_\lambda(Z)=UD^*V^T$.
+ Dla dostaecznie dużych wartości $\lambda$, $D^*$ will be rank-reduced, and hence so will be $UD^*V^T$.
 
+## Rozkład SVD
+
+Każdą macierz A można przedstawić jako:
+
+$$ A = USV $$
+gdzie:
+
+$U, V$ - macierze ortogolnalne
+
+$S$ - macierz diagonalna złożona z wartości singularnych $A$ 
 # Działanie

### 2020-03-31 raport.Rmd
algebra

@@
 ```{r}
-#require(deepdep)
-#plot_dependencies(deepdep("Amelia", depth = 2))
-#plot_dependencies(deepdep("missMDA", depth = 2))
-#plot_dependencies(deepdep("missForest", depth = 2))
-#plot_dependencies(deepdep("VIM", depth = 2))
-#plot_dependencies(deepdep("mice", depth = 2))
-#plot_dependencies(deepdep("softImpute", depth = 2))
+require(deepdep)
+plot_dependencies(deepdep("Amelia", depth = 2))
+plot_dependencies(deepdep("missMDA", depth = 2))
+plot_dependencies(deepdep("missForest", depth = 2))
+plot_dependencies(deepdep("VIM", depth = 2))
+plot_dependencies(deepdep("mice", depth = 2))
+plot_dependencies(deepdep("softImpute", depth = 2))
 ```
@@
 3. Tworzy nową macierz $Z$: $S_\lambda(Z)=UD^*V^T$.
- Dla dostaecznie dużych wartości $\lambda$, $D^*$ will be rank-reduced, and hence so will be $UD^*V^T$.
+Dla dostaecznie dużych wartości $\lambda$, $D^*$ będzie miało mniejszy rząd, więc rózwnież $UD^*V^T$.
 
@@
 $S$ - macierz diagonalna złożona z wartości singularnych $A$
+
 # Działanie

### 2020-04-04 raport.Rmd
algorytm

--- 2020-03-30 raport.Rmd
@@
 
-# Teoria
+# Jak działa algorytm?
 
-No dobra, ale czym właściwie była ta *trace norm*?
+## Soft-thresholded SVD
 
-Jest to suma wartości bezwzględnych wartości własnych macierzy,
+Niech $X$  będzie dużą $m\times n$ macierzą z wiloma brakującymi wartościami.
+Niech $\Omega$ zawiera pary indeksów $(i,j)$ gdzie $X$ był zaobserwowany, i niech $P_\Omega(X)$ oznacza macierz z wartościami jak w $X$ w punktach z $\Omega$, a w pozostałych ma wartości 0.
+Zatem jeśli $X$ ma brakujące wartości w $\Omega^\perp$, $P_\Omega(X)$ by oznaczyło NA jako 0.
 
-$||A||_{1} = trace((\sqrt{(X^{*}X)}))$
+Rozważmy wyrażenie
+$$\min_M\frac12\|P_\Omega(X)-M\|^2_F+\lambda\|M\|_*,$$
+gdzie $\|M\|_*$ to suma wartości singularnych macierzy $M$ (nucelar norm, trace norm), natomiast $\lambda$ jest parametrem.
+
+Niech $\widehat M$ będzie przybliżeniem rozwiązania problemu:
+$$Z=P_\Omega(X)+P_{\Omega^\perp}(\widehat M).$$
+
+$${\widehat M}=S_\lambda(Z)$$
+
+Operator $S_\lambda(Z)$ dla macierzy $Z$ wykonuje następujące instrukcje:
+
+1. Znajduje rozkład SVD dla $Z=UDV$, oraz $d_i$ - wartości singularne $Z$.
+2. Zaszumia wartości singularne: $d_i^*= (d_i-\lambda)_+$.
+3. Tworzy nową macierz $Z$: $S_\lambda(Z)=UD^*V$.
+Dla dostaecznie dużych wartości $\lambda$, $D^*$ będzie miało zmniejszony rząd, więc również i $UD^*V$.
+
+Stąd tworząc kolejne macierze ${\widehat M}$, dostajemy algorytm iteracyjny.
+
+## Rozkład SVD
+
+Każdą macierz A można przedstawić jako:
+
+$$ A = UDV $$
+gdzie:
+
+$U, V$ - macierze ortogolnalne
+
+$D$ - macierz diagonalna złożona z wartości singularnych $A$
 

### 2020-04-05 raport.Rmd
Jeszcze będę działał jutro

@@
 
-SoftImpute to nieduży i przydatny pakiet do uzupełniania macierzy, stworzoną przez Trevor Hastiego(Stanford, autor wielu książek o statystce).
+SoftImpute to nieduży i przydatny pakiet do uzupełniania macierzy, stworzony przez Trevor Hastiego (Stanford, autor wielu książek o statystce).
 Wykorzystuje do tego **nuclear norm regularization**, znanej także jako **trace norm regularization** (pl: *regularyzacja norm śladowych*).
@@
 
-  * biscale() - zoptymalizowane scale() działające również na niekompletnej macierzy, działające też szybko na bardzo dużych macierzach
+  * *biscale()* - zoptymalizowane scale() działające również na niekompletnej macierzy, działające też szybko na bardzo dużych macierzach
 
-  * dwie nowe klasy dla macierzy: SparseplusLowRank i Incomplete
+  * dwie nowe klasy dla macierzy: *SparseplusLowRank* i *Incomplete*
 
@@
 Pakiet z założenia służy do uzupełniania macierzy, wymaga on nas konwersji z data.frame na klasę matrix. Operuje tylko na macierzach, co jest równocześnie zaletą - potrzebuje tylko pakiet Matrix:
-```{r}
+```{r, warning = FALSE}
 require(deepdep)
@@
 
-## Przykład
+## Przykłady
 
-Wygenerujmy losową macierz 5 x 6 i usuńmy losowe 7 wartości.
+Wpierw w celu obycia ze składnią przetestujmy wpierw naszą podstawową funkcję na losowej macierzy 5 x 6 z wartościami z rozkładu normalnego i usuńmy z niej losowe 7 wartości.
 
 ```{r przyklad1_1}
-set.seed(123)
+set.seed(17)
 
 X = matrix(rnorm(30), 5, 6)
-X
 X_del = X
@@
 
-Zobaczmy wynik działania *soft_impute()*.
-W tym celu utworzymy zmienną *fits* generującą dopasowanie w zależności od macierzy z usuniętymi wartościami, a następnie wypełnimy ją z pomocą funkcji *complete()*.
+Utworzymy zmienną *fits* generującą dopasowanie w zależności od macierzy z usuniętymi wartościami, a następnie wypełnimy ją z pomocą funkcji *complete()*.
 
 ```{r przyklad1_2}
-fits = softImpute(X_del, trace = TRUE, type="svd")
-complete(X_del, fits)
+fits <- softImpute(X_del, trace = TRUE, type="svd")
+completed <- complete(X_del, fits)
 ```
 
-Jak widać funkcja poradziła sobie całkiem dobrze w przypadku losowo wygenerowanych wartości z rozkładu normalnego.
+Zobaczmy macierz wejściową, wypełnioną, ich różnicę i miarę rmse.
+
+```{r przyklad1_3}
+X
+completed
+X - completed
+# install.packages("Metrics") # if not installed
+library(Metrics)
+rmse(X, completed)
+```
+
+W kolejnych krokach będziemy rozpatrywać macierz wygenerowaną w bardziej złożony sposób zawierającą już pewne nie-zupełnie-losowe własności matematyczne. Wpierw przygotujmy ją i usuńmy 30% wartości.
+
+```{r przyklad2_1}
+set.seed(101)
+n <- 200
+p <- 100
+J <- 50
+np <- n * p
+missfrac <- 0.3
+
+x <- matrix(rnorm(n*J), n, J) %*% matrix(rnorm(J*p), J, p) + matrix(rnorm(np), n, p) / 5
+sample(x, 20)
+
+ix <- seq(np)
+imiss <- sample(ix, np*missfrac, replace=FALSE)
+xna <- x
+xna[imiss] = NA
+```
+
+Metoda "klasyczna"
+
+```{r przyklad2_2}
+### uses regular matrix method for matrices with NAs
+fit1 <- softImpute(xna, rank = 50, lambda = 30)
+```
+
+```{r przyklad2_3}
+### uses sparse matrix method for matrices of class "Incomplete"
+xnaC <- as(xna, "Incomplete")
+fit2 <- softImpute(xnaC, rank=50, lambda=30)
+```
+
+```{r przyklad2_4}
+### uses "svd" algorithm
+fit3 <- softImpute(xnaC, rank=50, lambda=30, type="svd")
+ximp <- complete(xna,fit1)
+```
+
+```{r przyklad2_5}
+### first scale xna
+xnas <- biScale(xna)
+fit4 <- softImpute(xnas,rank=50,lambda=10)
+ximp <- complete(xna,fit4)
+impute(fit4, i = c(1, 3, 7), j = c(2, 5, 10))
+impute(fit4, i = c(1, 3, 7), j = c(2, 5, 10), unscale=FALSE) #ignore scaling and centering
+```
 

### 2020-04-06 raport.Rmd
Chyba ostateczna

@@
 
-SoftImpute to nieduży i przydatny pakiet do uzupełniania macierzy, stworzony przez Trevor Hastiego (Stanford, autor wielu książek o statystce).
+SoftImpute to nieduży i przydatny pakiet do uzupełniania macierzy, stworzony przez Trevor Hastiego (Stanford, autor wielu książek o statystyce).
 Wykorzystuje do tego **nuclear norm regularization**, znanej także jako **trace norm regularization** (pl: *regularyzacja norm śladowych*).
@@
 
-    2. Uzupełnianie braków ramki szacunkami korzystając z alternating ridge regression(zwane popularniej w Europie Tikhonov regularization) (opcja *type="als"*)
+    2. Uzupełnianie braków ramki szacunkami korzystając z alternating ridge regression (zwane popularniej w Europie *Tikhonov regularization*) (opcja *type="als"*)
 
@@
 
+### Ogląd na prostej macierzy
+
 Wpierw w celu obycia ze składnią przetestujmy wpierw naszą podstawową funkcję na losowej macierzy 5 x 6 z wartościami z rozkładu normalnego i usuńmy z niej losowe 7 wartości.
@@
 ```{r przyklad1_2}
-fits <- softImpute(X_del, trace = TRUE, type="svd")
+fits <- softImpute(X_del, trace = TRUE, type = "svd")
 completed <- complete(X_del, fits)
@@
 
-```{r przyklad1_3}
+```{r przyklad1_3, warning = FALSE}
 X
@@
 ```
+
+### Metoda regularna
 
@@
 
-```{r przyklad2_3}
+W efekcie otrzymujemy 3 macierze:
+
+1. fit1\$u odpowiada za $U$
+2. fit1\$v odpowiada za $V$
+3. fit1\$d odpowiada za $\Sigma$
+
+...gdzie:
+
+* ${\displaystyle A=U\Sigma V^{T},}$
+
+* $U$ i $V$ – macierze ortogonalne (czyli ${\displaystyle U^{-1}=U^{T},}{\displaystyle U^{-1}=U^{T},} {\displaystyle V^{-1}=V^{T}}V^{-1}=V^T)$
+
+* $\Sigma$  – macierz diagonalna (przekątniowa), taka że ${\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}{\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}$ gdzie$ {\displaystyle \sigma _{i}}\sigma _{i}$ – nieujemne wartości szczególne (osobliwe) macierzy ${\displaystyle A,}A$, zwyczajowo uporządkowane nierosnąco.
+
+```{r przyklad2_3, echo = FALSE}
+print("Wymiary V:")
+print(dim(fit1$v))
+print("Wymiary U:")
+print(dim(fit1$u))
+print("Wartości macierzy diagonalnej:")
+print(fit1$d)
+```
+
+Zobaczmy czy zaiste wygenerowane macierze spełniają warunek rozkładu SVD i średni błąd kwadratowy:
+
+```{r przyklad2_4}
+result <- (fit1$u * fit1$d) %*% t(fit1$v)
+rmse(result, x)
+```
+
+Satysfakcjonująco! A jak poradzi sobie funkcja *complete()*?
+
+```{r przyklad2_5}
+completed <- complete(xna, fit1)
+rms1 <- rmse(x, completed)
+rms1
+```
+
+30% brakujących wartości zostało uzupełnionych i wynik wyszedł jeszcze lepszy.
+
+### Dla klasy *Incomplete*
+
+W dalszej kolejności efekt działania klasy *Incomplete* z użytkiem macierzy rzadkich.
+
+```{r przyklad3_1}
 ### uses sparse matrix method for matrices of class "Incomplete"
@@
 
-```{r przyklad2_4}
+```{r przyklad3_2, echo = FALSE}
+print("Wymiary V:")
+print(dim(fit2$v))
+print("Wymiary U:")
+print(dim(fit2$u))
+```
+
+```{r przyklad3_3}
+result <- (fit2$u * fit2$d) %*% t(fit2$v)
+rmse(result, x)
+completed <- complete(xna, fit2)
+rms2 <- rmse(x, completed)
+rms2
+```
+
+### Algorytm *svd*
+
+... czyli iteracyjne obliczanie *soft-thresholded SVD* z wypełnionej macierzy.
+
+```{r przyklad4}
 ### uses "svd" algorithm
 fit3 <- softImpute(xnaC, rank=50, lambda=30, type="svd")
-ximp <- complete(xna,fit1)
+
+result <- (fit3$u * fit3$d) %*% t(fit3$v)
+rmse(result, x)
+completed <- complete(xna, fit3)
+rms3 <- rmse(x, completed)
+rms3
 ```
 
-```{r przyklad2_5}
+### biScale()
+
+Całkiem ładny wynik. Zobaczmy jeszcze *softImpute()* z użyciem *biScale()*.
+
+```{r przyklad5_1}
 ### first scale xna
 xnas <- biScale(xna)
-fit4 <- softImpute(xnas,rank=50,lambda=10)
-ximp <- complete(xna,fit4)
+fit4 <- softImpute(xnas, rank=50, lambda=10)
+completed <- complete(xna, fit4)
+rms4 <- rmse(x, completed)
+rms4
+```
+
+### impute() pod lupą
+
+Możemy też przyjrzeć się lepiej temu co siedzi pod *complete()* wyświetlając efekt samej funkcji *impute()*, z której *complete()* korzysta.
+
+W dokumentacji czytamy: *impute returns a vector of predictions, using the reconstructed low-rank matrix representation represented by object.
+It is used by complete, which returns a complete matrix with all the missing values imputed.*
+
+```{r przyklad6}
 impute(fit4, i = c(1, 3, 7), j = c(2, 5, 10))
-impute(fit4, i = c(1, 3, 7), j = c(2, 5, 10), unscale=FALSE) #ignore scaling and centering
+impute(fit4, i = c(1, 3, 7), j = c(2, 5, 10), unscale = FALSE) # ignore scaling and centering
+```
+
+### Podsumowanie
+
+Na dobry koniec porównajmy jeszcze miary błędów średniokwadratowych dla zaproponowanych metod imputacji macierzy:
+
+```{r przyklady_podsumowanie}
+summary <- data.frame(rms1, rms2, rms3, rms4)
+colnames(summary) <- c("regular", "macierze rzadkie + incomplete", "svd", "z bi-skalowaniem")
+summary
 ```

### 2020-04-06 raport.Rmd
Poprawiona część Kuby po próbie

@@
 
+UWAGA! Tylko wartości numeryczne.
+
+```{r przyklad0, eval = FALSE}
+softImpute(matrix(c('a', 'b', 'a', NA, 'c', 'b'), 2, 3))
+```
+
+Wywołanie powyższej linii kodu zwróciłoby błąd.
+
 ### Ogląd na prostej macierzy
@@
 
+Przygotujemy także dwie maierze pomocnicze - *xMeans* i *xZeros* odpowiadające za matryce uzupełnione odpowiedni średnimi i zerami.
+Porównanie miar rezultatu *softImpute()* i owych dwóch da nam dobry ogląd jakości naszej funkcji.
+
+```{r przyklad2_2}
+xZeros <- xna
+xZeros[is.na(xZeros)] <- 0
+sample(xZeros, 10)
+
+Xmean <- mean(xZeros) * 10 / 7
+print(Xmean)
+xMeans <- xna
+xMeans[is.na(xMeans)] <- Xmean
+sample(xMeans, 10)
+```
+
 Metoda "klasyczna"
 
-```{r przyklad2_2}
+```{r przyklad2_3}
 ### uses regular matrix method for matrices with NAs
@@
 
-* $\Sigma$  – macierz diagonalna (przekątniowa), taka że ${\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}{\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}$ gdzie$ {\displaystyle \sigma _{i}}\sigma _{i}$ – nieujemne wartości szczególne (osobliwe) macierzy ${\displaystyle A,}A$, zwyczajowo uporządkowane nierosnąco.
+* $\Sigma$  – macierz diagonalna (przekątniowa), taka że ${\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}{\displaystyle \Sigma =\operatorname {diag} (\sigma _{i}),}$ gdzie ${\displaystyle \sigma _{i}}\sigma _{i}$ – nieujemne wartości szczególne (osobliwe) macierzy ${\displaystyle A,}A$, zwyczajowo uporządkowane nierosnąco.
 
-```{r przyklad2_3, echo = FALSE}
+```{r przyklad2_4, echo = FALSE}
 print("Wymiary V:")
@@
 
-Zobaczmy czy zaiste wygenerowane macierze spełniają warunek rozkładu SVD i średni błąd kwadratowy:
+Zobaczmy czy zaiste wygenerowane macierze spełniają warunek rozkładu SVD i średni błąd kwadratowy macierzy wejściowej i wygenerowanej przez *softImpute()*:
 
-```{r przyklad2_4}
+```{r przyklad2_5}
 result <- (fit1$u * fit1$d) %*% t(fit1$v)
 rmse(result, x)
+rmsM <- rmse(x, xMeans)
+rmsM
+rms0 <- rmse(x, xZeros)
+rms0
 ```
+
+W porównaniu do macierzy wejściowej uzupełnionej zerami i średnimi tak średnio bym powiedział.
+Ale w końcu robotę robi funkcja *complete()* - która uzupełnia braki macierzy wejściowej tymi z tej specjalnie generowanej.
+Oczywiście im mniejszy błąd - tym lepszy będzie rezultat.
+Ale czy aż dla 30% funkcja ogarnie coś sensownego?
 
@@
 
-```{r przyklad2_5}
+```{r przyklad2_6}
 completed <- complete(xna, fit1)
@@
 
-30% brakujących wartości zostało uzupełnionych i wynik wyszedł jeszcze lepszy.
+Jak widać jak najbardziej! 30% brakujących wartości zostało uzupełnionych i wynik wyszedł zdecydowanie lepszy niż dla uzupełniania zerami czy średnimi.
 
@@
 xnas <- biScale(xna)
+```
+
+Czy faktycznie nasza macierz została poprawnie przeskalowana?
+
+```{r przyklad5_2}
+mean(xnas[!is.na(xnas)])
+var(xnas[!is.na(xnas)])
+```
+
+Wszystko jest w porządku! Średnia wyszła niemalże 0, a wariancja - niemalże 1. Otrzymaliśmy więc poprawne przeskalowanie.
+
+```{r przyklad5_3}
 fit4 <- softImpute(xnas, rank=50, lambda=10)
@@
 summary
+
+zerosMeans <- data.frame(rms0, rmsM)
+colnames(zerosMeans) <- c("zeros", "means")
+zerosMeans
 ```
+
+Sukces!
 

### 2020-04-19 szkielet.r
fixed get_result + minor changes

@@ -9,13 +9,13 @@ library(VIM)
 # not needed directly, but have to be installed
 #library(jsonlite)
 
-DFT_REPO_DATASET_DI? = './dependencies/datasets'
+DFT_REPO_DATASET_DIR = './dependencies/datasets'
 getwd()
 
 #' Reads dataset based on id from directory
 #' @param openml_id int. Should be present in dataset_dir
 #' @param dataset_dir string with directory with subdirectories like 'openml_dataset_<openml_id>'
-#' @return list, w?ere list$dataset contains processed data
+#' @return list, where list$dataset contains processed data
 #' @seealso names(list)
 read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
   
@@ -23,7 +23,7 @@ read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
     stop(paste(dataset_dir, 'does not exist' ))
   }
   
-  dir <- paste(dataset_dir,?paste('openml_dataset', openml_id, sep = '_'), sep ='/')
+  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
   if (!dir.exists(dir)){
     stop(paste(dir, 'does not exist' ))
   }
@@ -32,7 +32,7 @@ read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
   
   #set right dir to code.R to acually work - it depends on dirlocation to create json
   setwd(dir)
-  ?  # use new env to avoid trashing globalenv
+   # use new env to avoid trashing globalenv
   surogate_env = new.env(parent = .BaseNamespaceEnv)
   attach(surogate_env)
   source("code.R",surogate_env)
@@ -40,7 +40,7 @@ read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
   j <- jsonlite::read_json('./dataset.json')
   j$dataset <- surogate_env$dataset
   setwd(start_dir)
- ?
+ 
   return(j)
 }
 
@@ -61,7 +61,7 @@ read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
 #' cat('Number of features with missing data: ', n)
 #' 
 #' @seealso names(list)
-read_all_datasets <- function(dat?set_dir = DFT_REPO_DATASET_DIR){
[…]

### 2020-04-20 parallel.r
parallelelele

@@ -44,7 +44,7 @@ library(batchtools)
 
 # ułatwienie: btlapply
 
-registry <- makeRegistry(file.dir = "./file_registry", seed = 15390)
+registry <- makeRegistry(file.dir = "./registry", seed = 15390)
 getDefaultRegistry()
 # setDefaultRegistry()
 # clearRegistry()

### 2020-04-20 szkielet.r

@@ -182,7 +182,11 @@ f1 <- function(confusion_matrix){
 
 
 library(rpart) # for classification tree
-get_result <- function(dataset, imputation_fun, name_of_target){
+get_result <- function(dataset_list, imputation_fun){
+  
+  dataset <- dataset_list$dataset
+  name_of_target <- dataset_list$target
+  
   # imputation
   imputation_start = Sys.time() # start to measure time
   imputated_dataset <- imputation_fun(dataset) 
@@ -215,36 +219,68 @@ get_result <- function(dataset, imputation_fun, name_of_target){
   colnames(classification_report) <- c("accuracy", "precision",
                                        "recall", "f1")
   
-  return(list(confusion_matrix = confusion_matrix,
-                    classification_report = classification_report,
-                    imputation_time = imputation_stop - imputation_start,
-                    modelling_time = modelling_stop - modelling_start))
+  dataset_list$dataset <- NULL
+  # in future maybe return all dataset_list ?
+  # for now stick with readability
+  
+  imp_method_name <- as.character(substitute(imputation_fun))
+  
+  return(list( dataset_id = dataset_list$id, 
+               imp_method = imp_method_name,
+               confusion_matrix = confusion_matrix,
+               classification_report = classification_report,
+               imputation_time = imputation_stop - imputation_start,
+               modelling_time = modelling_stop - modelling_start))
 }
-# example: get_result(data_all[[2]]$dataset, imputation_fun_vim, data_all[[2]]$target)
+# example: get_result(data_all[[1]], imputation_remove_rows)
 # TODO: dokumentacja
+
+
 
 # reading datasets from OpenML's edits
 data_all <- read_all_datasets()
 
-
-#PARALLEL ROBI BRRRRR [TODO bo nie wiem jak]
-# ...
-# ...
-# ...
-
-
 # imputations and targets preparation
 imputations <- list(imputation_fun_vim, imputation_fun_missForest,
                     imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
-targets <- list()
-for (i in 1:length(data_all)){
-  targets[[i]] <- data_all[[i]]$target
-}
+
+targets <- lapply(data_all, function(d){d$target})
+
+
+#PARALLEL ROBI BRRRRR [TODO bo nie wiem jak]
+library(batchtools)
+
+loadRegistry(file.dir = './registry', writeable = TRUE)
+
[…]

### 2020-04-21 reading_datasets_functions.r
move functions to files

@@ -1,0 +1,66 @@
+DFT_REPO_DATASET_DIR = './dependencies/datasets'
+getwd()
+
+#' Reads dataset based on id from directory
+#' @param openml_id int. Should be present in dataset_dir
+#' @param dataset_dir string with directory with subdirectories like 'openml_dataset_<openml_id>'
+#' @return list, where list$dataset contains processed data
+#' @seealso names(list)
+read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
+  
+  if (!dir.exists(dataset_dir)){
+    stop(paste(dataset_dir, 'does not exist' ))
+  }
+  
+  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
+  if (!dir.exists(dir)){
+    stop(paste(dir, 'does not exist' ))
+  }
+  
+  start_dir <- getwd()
+  
+  #set right dir to code.R to acually work - it depends on dirlocation to create json
+  setwd(dir)
+  # use new env to avoid trashing globalenv
+  surogate_env <- new.env(parent = .BaseNamespaceEnv)
+  attach(surogate_env)
+  source("code.R",surogate_env)
+  
+  j <- jsonlite::read_json('./dataset.json')
+  j$dataset <- surogate_env$dataset
+  setwd(start_dir)
+  
+  return(j)
+}
+
+
+#' Reads all dataset from given directory
+#' @param dataset_dir string with directory with subdirectories like 'openml_dataset_<openml_id>'
+#' 
+#' @return matrix with columns = c(dataset, target_name)
+#' 
+#' @example 
+#' dfs <- read_all_da?asets()
+#' df <- dfs[[1]]
+#' 
+#' target_name <- df$target
+#' cat('Target name: ', target_name)
+#' 
+#' n <- df$number_of_features_with_missings
+#' cat('Number of features with missing data: ', n)
+#' 
+#' @seealso names(list)
+read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){
+  
+  if (!dir.exists(dataset_dir)){
+    stop(paste(dataset_dir, 'does not exist'))
+  }
+  
+  start_dir <- getwd()
+  subdirs <- dir(dataset_dir)
+  ids <- sapply(subdirs, function(dir){substr(dir, 16, nchar(dir))})
+  datasets_combined <- lapply(ids, function(x){read_dataset(x, dataset_dir)})
[…]

### 2020-04-21 szkielet.r

@@ -2,184 +2,9 @@
 #getwd()
 set.seed(123)
 
-library(mice)
-library(missForest)
-library(VIM)
-
-# not needed directly, but have to be installed
-#library(jsonlite)
-
-DFT_REPO_DATASET_DIR = './dependencies/datasets'
-getwd()
-
-#' Reads dataset based on id from directory
-#' @param openml_id int. Should be present in dataset_dir
-#' @param dataset_dir string with directory with subdirectories like 'openml_dataset_<openml_id>'
-#' @return list, where list$dataset contains processed data
-#' @seealso names(list)
-read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
-  
-  if (!dir.exists(dataset_dir)){
-    stop(paste(dataset_dir, 'does not exist' ))
-  }
-  
-  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
-  if (!dir.exists(dir)){
-    stop(paste(dir, 'does not exist' ))
-  }
-  
-  start_dir <- getwd()
-  
-  #set right dir to code.R to acually work - it depends on dirlocation to create json
-  setwd(dir)
-   # use new env to avoid trashing globalenv
-  surogate_env = new.env(parent = .BaseNamespaceEnv)
-  attach(surogate_env)
-  source("code.R",surogate_env)
-  
-  j <- jsonlite::read_json('./dataset.json')
-  j$dataset <- surogate_env$dataset
-  setwd(start_dir)
- 
-  return(j)
-}
-
-
-#' Reads all dataset from given directory
-#' @param dataset_dir string with directory with subdirectories like 'openml_dataset_<openml_id>'
-#' 
-#' @return matrix with columns = c(dataset, target_name)
-#' 
-#' @example 
-#' dfs <- read_all_da?asets()
-#' df <- dfs[[1]]
-#' 
-#' target_name <- df$target
-#' cat('Target name: ', target_name)
-#' 
-#' n <- df$number_of_features_with_missings
-#' cat('Number of features with missing data: ', n)
-#' 
-#' @seealso names(list)
-read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){
-  
-  if (!dir.exists(dataset_dir)){
-    stop(paste(dataset_dir, 'does not exist'))
-  }
-  
-  start_dir <- getwd()
-  subdirs <- dir(dataset_dir)
[…]

### 2020-04-22 pd2.Rmd
Szkic raportu, jeszcze bez parallela

---
title: "The imputation methods big test"
author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
date: "22/04/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Choosing the best-suited imputation is the daily dilemma of every data scientist. Some of they believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.

In this document we will try to find one objectively the finest imputation method by testing five of popular ones on eight specially selected data sets prepared for these operations.

## Imputation functions

We'll look at the two simple imputation methods with the following short markings:

1. *I1* / *modeMedian* - missing data supplemented with medians from individual columns
2. *I2* / *removeRows* - all the rows with any nonexistent values removed

```{r imputation_basic_functions}
imputation_mode_median <- function(df){

  Mode <- function(x) {
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
  }

  for (i in 1L:length(df)){
    if (sum(is.na(df[,i])) > 0){
      if (mode(df[,i]) == 'character' | is.factor(df[,i])){
        to_imp <- Mode(df[,i])
        df[,i][is.na(df[,i])] <- to_imp
      }
      else{
        to_imp <- median(df[,i], na.rm = TRUE) 
        df[,i][is.na(df[,i])] <- to_imp
      }
    }
  }

  return(df)
}

imputation_remove_rows <- function(df){
  return (na.omit(df))
}
```

... and also three more advanced algorithms from popular libraries:

3. *I3* - *mice*
4. *I4* - *vim*
5. *I5* - *missForest*

```{r imputation_advanced_functions, message = FALSE, warning = FALSE}
library(mice)
imputation_fun_mice <- function(df){
  init <- mice(df, maxit=0) 
  meth <- init$method
  predM <- init$predictorMatrix
  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5)
  completed <- complete(imputed)
  return(completed)
}

library(VIM)
imputation_fun_vim <- function(df){
  no_columns <- length(df)
  imputed <- kNN(df)
  imputed <- imputed[,1:no_columns]
  return(imputed)
}

library(missForest)
imputation_fun_missForest <- function(df){
  return(missForest(df)$ximp)
}
```

## Reading datasets

The eight data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:

* 29 - *credit-approval*
* 38 - *sick*
* 188 - *eucalyptus*
* 1590 - *adult*
* 6332 - *cylinder-bands*
* 23381 - *dresses-sales*
* 40536 - *SpeedDating*
* 41278 - *okcupid-stem*

Those above have been placed in individual directories identified by id in the prepared directory.

```{r read_dataset}
DFT_REPO_DATASET_DIR = './dependencies/datasets'

read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist' ))
  }

  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
  if (!dir.exists(dir)){
    stop(paste(dir, 'does not exist' ))
  }

  start_dir <- getwd()

  # set right dir to code.R to acually work - it depends on dirlocation to create json
  setwd(dir)
  # use new env to avoid trashing globalenv
  surogate_env <- new.env(parent = .BaseNamespaceEnv)
  attach(surogate_env)
  source("code.R",surogate_env)

  j <- jsonlite::read_json('./dataset.json')
  j$dataset <- surogate_env$dataset
  setwd(start_dir)

  return(j)
}
```

In order to read all the eight datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.

```{r read_all_datasets}
read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist'))
  }

  start_dir <- getwd()
  subdirs <- dir(dataset_dir)
  ids <- sapply(subdirs, function(dir){substr(dir, 16, nchar(dir))})
  datasets_combined <- lapply(ids, function(x){read_dataset(x, dataset_dir)})

  datasets_combined <- t(datasets_combined)
  return(unname(t(datasets_combined)))
}
```

## Metrics functions

In order to test the same test-train splits we'll use one random seed 1357 for all datasets.

```{r train_test_split}
set.seed(1357)

train_test_split <- function(dataset, train_size){
  smp_size <- floor(train_size * nrow(dataset))
  typeof(smp_size)

  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)

  train <- dataset[train_ind, ]
  test <- dataset[-train_ind, ]

  return (list(train, test))
}
```

For each machine learning model after every imputation we will get the confusion matrix and the values of four basic metrics:

* *accuracy* - $\frac{TP+TN}{TP+FP+FN+TN}$
* *precision* - $\frac{TP}{TP+FP}$
* *recall* - $\frac{TP}{TP+FN}$
* *f1* - $2*\frac{Recall * Precision}{Recall + Precision}$

```{r metrics}
get_confusion_matrix <- function(test, pred){
  return (table(Truth = test, Prediction = pred))
}

confusion_matrix_values <- function(confusion_matrix){
  TP <- confusion_matrix[2,2]
  TN <- confusion_matrix[1,1]
  FP <- confusion_matrix[1,2]
  FN <- confusion_matrix[2,1]
  return (c(TP, TN, FP, FN))
}

accuracy <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return((conf_matrix[1] + conf_matrix[2]) / (conf_matrix[1] + conf_matrix[2] + conf_matrix[3] + conf_matrix[4]))
}

precision <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1]/ (conf_matrix[1] + conf_matrix[3]))
}

recall <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1] / (conf_matrix[1] + conf_matrix[4]))
}

f1 <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  rec <- recall(confusion_matrix)
  prec <- precision(confusion_matrix)
  return(2 * (rec * prec) / (rec + prec))
}
```

## Imputation results

```{r ger_result_function}
library(rpart)
get_result <- function(dataset_list, imputation_fun){

  dataset <- dataset_list$dataset
  name_of_target <- dataset_list$target

  # imputation
  imputation_start = Sys.time() # start to measure time
  imputated_dataset <- imputation_fun(dataset) 
  imputation_stop = Sys.time() # end measuring time

  # train test split
  train_test <- train_test_split(imputated_dataset, 0.8)
  train <- as.data.table(train_test[1])
  test <- as.data.table(train_test[2])

  # modelling
  vars <- colnames(dataset)[colnames(dataset)!=name_of_target]
  my_formula <- as.formula(paste(name_of_target, paste(vars, collapse=" + "), sep=" ~ "))
  modelling_start = Sys.time() # start to measure time
  tree_model <- rpart(formula = my_formula, data = train,
                      method = "class", control = rpart.control(cp = 0))
  y_pred <- as.data.frame(predict(tree_model, test, type = "class"))
  modelling_stop = Sys.time() # end measuring time

  # calculating metrics
  confusion_matrix <- get_confusion_matrix(test[[name_of_target]], y_pred[,1])

  accuracy_v <- accuracy(confusion_matrix)
  precision_v <- precision(confusion_matrix)
  recall_v <- recall(confusion_matrix)
  f1_v <- f1(confusion_matrix)

  classification_report <- data.frame(accuracy_v, precision_v,
                                      recall_v, f1_v)
  colnames(classification_report) <- c("accuracy", "precision",
                                       "recall", "f1")

  dataset_list$dataset <- NULL
  # in future maybe return all dataset_list ?
  # for now stick with readability

  imp_method_name <- deparse(substitute(imputation_fun))

  return(list( dataset_id = dataset_list$id, 
               imp_method = imp_method_name,
               confusion_matrix = confusion_matrix,
               classification_report = classification_report,
               imputation_time = imputation_stop - imputation_start,
               modelling_time = modelling_stop - modelling_start))
}

data_all <- read_all_datasets()

# imputations and targets preparation
imputations <- list(imputation_fun_vim, imputation_fun_missForest,
                    imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
targets <- lapply(data_all, function(d){d$target})
```

Here are the results:

```{r print_results}
for (imputation in imputations){
  for(i in 1:8){
    results <- get_result(data_all[[i]], imputation)
    print(paste0("Dataset id: ", results$id, ", method: ", results$imp_method))
    print(paste0("Confusion matrix: ", results$confusion_matrix))
    print(paste0("Classification report: ", results$classification_report))
    print(paste0("Imputation time: ", results$imputation_time))
    print(paste0("Modelling time: ", results$modelling_time))
  }
}
```

### 2020-04-22 reading_datasets_functions.r
Jakaś tam poprawka, chyba dokumentacji

@@ -40,7 +40,7 @@ read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
 #' @return matrix with columns = c(dataset, target_name)
 #' 
 #' @example 
-#' dfs <- read_all_da?asets()
+#' dfs <- read_all_datasets()
 #' df <- dfs[[1]]
 #' 
 #' target_name <- df$target

### 2020-04-22 szkielet.r
parallele w miare dzialaja

@@ -72,6 +72,42 @@ imputations <- list(imputation_fun_vim, imputation_fun_missForest,
 targets <- lapply(data_all, function(d){d$target})
 
 
+
+#PARALLEL ROBI BRRRRR I TEN JAKO JEDYNY DZIAŁA!
+library(parallelMap)
+parallelStartMulticore(8, show.info = TRUE)
+
+res1 <- parallelLapply(data_all, function(x){get_result(x,imputation_remove_rows)}, impute.error = function(x){'ERROR'})
+res2 <- parallelLapply(data_all, function(x){get_result(x,imputation_mode_median)}, impute.error = function(x){'ERROR'})
+res3 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_mice)}, impute.error = function(x){'ERROR'})
+res4 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_missForest)}, impute.error = function(x){'ERROR'})
+res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
+
+
+
+
+### Tu juz nie dziala
+datasets_all <- lapply(data_all, function(x){x$dataset})
+res <- parallelLapply(datasets_all, imputation_fun_mice)
+
+library(parallel)
+
+parallel::setDefaultCluster()
+Job1 = mcparallel(res <- parallelLapply(datasets_all, imputation_fun_mice, impute.error = function(x){'ERROR'} ))
+Job2 = mcparallel(res <- parallelLapply(datasets_all, imputation_remove_rows, impute.error = function(x){'ERROR'}))
+Job3 = mcparallel(res <- parallelLapply(datasets_all, imputation_mode_median, impute.error = function(x){'ERROR'}))
+Job4 = mcparallel(res <- parallelLapply(datasets_all, imputation_fun_vim, impute.error = function(x){'ERROR'}))
+Job5 = mcparallel(res <- parallelLapply(datasets_all, imputation_fun_missForest, impute.error = function(x){'ERROR'}))
+
+(res <- parallelLapply(data_all, function(x){get_result(x,imputation_remove_rows)}, impute.error = function(x){'ERROR'}))
+
+
+JobResult1 = mccollect(Job1)
+JobResult2 = mccollect(Job2)
+JobResult3 = mccollect(Job3)
+JobResult4 = mccollect(Job4)
[…]

### 2020-04-24 pd2.Rmd
tabelki robią brrrr

@@
 
-```{r ger_result_function}
+```{r get_result_function}
 library(rpart)
@@
 
-```{r print_results}
-for (imputation in imputations){
-  for(i in 1:8){
-    results <- get_result(data_all[[i]], imputation)
-    print(paste0("Dataset id: ", results$id, ", method: ", results$imp_method))
-    print(paste0("Confusion matrix: ", results$confusion_matrix))
-    print(paste0("Classification report: ", results$classification_report))
-    print(paste0("Imputation time: ", results$imputation_time))
-    print(paste0("Modelling time: ", results$modelling_time))
+```{r warning=FALSE}
+
+  library(knitr)
+  results <- c(list(0),list(0),list(0),list(0),list(0))
+  results[[1]] <- readRDS('./part_results/res1.rds')
+  results[[2]] <- readRDS('./part_results/res2.rds')
+  results[[3]] <- readRDS('./part_results/res3.rds')
+  results[[4]] <- readRDS('./part_results/res4.rds')
+  results[[5]] <- readRDS('./part_results/res5.rds')
+
+  result_table <- function(ds_it){
+    d <- data.frame(matrix(ncol = 7, nrow = 0))
+    colnames(d) <- c("imputation", "accuracy", "precision", "recall", "f1", "imp_time", "mod_time")
+    for (i in 1:length(imputations)){
+      if (length(results[[i]][[ds_it]]) != 1){ # length equal 1 means "ERROR" was generated in partial results 
+        d[i,] <- c(results[[i]][[ds_it]]$imp_method,
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$accuracy),3),   
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$precision),3),
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$recall),3),
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$f1),3),
+               round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
+               round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
       }
     }
+    d
+}
+
+  # change i bound to number of datasets 
+  for (i in c(1,2,3,5,6,7,8,9)){
+    kable(result_table(i), caption = paste("Dataset id: ", results[[1]][[i]]$dataset_id, sep = ""))
+  }
+
+  # jest taki problem że kable w pętli nie wypluwa wyniku, chyba że damy print(kable()) wtedy jednak brzydko wygląda
+
+  kable(result_table(1), caption = paste("Dataset id: ", results[[1]][[1]]$dataset_id, sep = ""))
+  kable(result_table(2), caption = paste("Dataset id: ", results[[1]][[2]]$dataset_id, sep = ""))
+  kable(result_table(3), caption = paste("Dataset id: ", results[[1]][[3]]$dataset_id, sep = ""))
+  kable(result_table(5), caption = paste("Dataset id: ", results[[1]][[5]]$dataset_id, sep = ""))
+  kable(result_table(6), caption = paste("Dataset id: ", results[[1]][[6]]$dataset_id, sep = ""))
+  kable(result_table(7), caption = paste("Dataset id: ", results[[1]][[7]]$dataset_id, sep = ""))
+  kable(result_table(8), caption = paste("Dataset id: ", results[[1]][[8]]$dataset_id, sep = ""))
+  kable(result_table(9), caption = paste("Dataset id: ", results[[1]][[9]]$dataset_id, sep = ""))
+
 ```
+
+```{r print_results}
+# for (imputation in imputations){
+#  for(i in 1:8){
+#    results <- get_result(data_all[[i]], imputation)
+#    print(paste0("Dataset id: ", results$id, ", method: ", results$imp_method))
+#    print(paste0("Confusion matrix: ", results$confusion_matrix))
+#    print(paste0("Classification report: ", results$classification_report))
+#    print(paste0("Imputation time: ", results$imputation_time))
+#    print(paste0("Modelling time: ", results$modelling_time))
+#  }
+#}
+```

### 2020-04-25 pd2.Rmd
Szkic raportu ostatecznego m. in.

--- 2020-04-22 pd2.Rmd
@@
 }
-
 imputation_remove_rows <- function(df){
@@
 }
-
 library(VIM)
@@
 }
-
 library(missForest)
@@
 
-* 29 - *credit-approval*
+* 29 - *credit-approval* PÓKI CO NIE BANGLA
 * 38 - *sick*
+* 56 - ...
 * 188 - *eucalyptus*
@@
 DFT_REPO_DATASET_DIR = './dependencies/datasets'
-
 read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
@@
 set.seed(1357)
-
 train_test_split <- function(dataset, train_size){
@@
 }
-
 confusion_matrix_values <- function(confusion_matrix){
@@
 }
-
 accuracy <- function(confusion_matrix){
@@
 }
-
 precision <- function(confusion_matrix){
@@
 }
-
 recall <- function(confusion_matrix){
@@
 }
-
 f1 <- function(confusion_matrix){
@@
 
-```{r ger_result_function}
+```{r get_result_function, warning = FALSE, message = FALSE}
 library(rpart)
@@
 }
-
 data_all <- read_all_datasets()
-
 # imputations and targets preparation
@@
 
-```{r print_results}
-for (imputation in imputations){
-  for(i in 1:8){
-    results <- get_result(data_all[[i]], imputation)
-    print(paste0("Dataset id: ", results$id, ", method: ", results$imp_method))
-    print(paste0("Confusion matrix: ", results$confusion_matrix))
-    print(paste0("Classification report: ", results$classification_report))
-    print(paste0("Imputation time: ", results$imputation_time))
-    print(paste0("Modelling time: ", results$modelling_time))
+```{r warning=FALSE}
+  library(knitr)
+  results <- c(list(0),list(0),list(0),list(0),list(0))
+  results[[1]] <- readRDS('./part_results/res1.rds')
+  results[[2]] <- readRDS('./part_results/res2.rds')
+  results[[3]] <- readRDS('./part_results/res3.rds')
+  results[[4]] <- readRDS('./part_results/res4.rds')
+  results[[5]] <- readRDS('./part_results/res5.rds')
+  result_table <- function(ds_it){
+    d <- data.frame(matrix(ncol = 7, nrow = 0))
+    colnames(d) <- c("imputation", "accuracy", "precision", "recall", "f1", "imp_time", "mod_time")
+    for (i in 1:length(imputations)){
+      if (length(results[[i]][[ds_it]]) != 1){ # length equal 1 means "ERROR" was generated in partial results 
+        d[i,] <- c(results[[i]][[ds_it]]$imp_method,
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$accuracy),3),   
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$precision),3),
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$recall),3),
+               round(as.numeric(results[[i]][[ds_it]]$classification_report$f1),3),
+               round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
+               round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
       }
     }
+    d
+}
+  # change i bound to number of datasets 
+  for (i in c(1,2,3,5,6,7,8,9)){
+    kable(result_table(i), caption = paste("Dataset id: ", results[[1]][[i]]$dataset_id, sep = ""))
+  }
+
+  # jest taki problem że kable w pętli nie wypluwa wyniku, chyba że damy print(kable()) wtedy jednak brzydko wygląda
+
+  kable(result_table(1), caption = paste("Dataset id: ", results[[1]][[1]]$dataset_id, sep = ""))
+  kable(result_table(2), caption = paste("Dataset id: ", results[[1]][[2]]$dataset_id, sep = ""))
+  kable(result_table(3), caption = paste("Dataset id: ", results[[1]][[3]]$dataset_id, sep = ""))
+  kable(result_table(5), caption = paste("Dataset id: ", results[[1]][[5]]$dataset_id, sep = ""))
+  kable(result_table(6), caption = paste("Dataset id: ", results[[1]][[6]]$dataset_id, sep = ""))
+  kable(result_table(7), caption = paste("Dataset id: ", results[[1]][[7]]$dataset_id, sep = ""))
+  kable(result_table(8), caption = paste("Dataset id: ", results[[1]][[8]]$dataset_id, sep = ""))
+  kable(result_table(9), caption = paste("Dataset id: ", results[[1]][[9]]$dataset_id, sep = ""))
+
 ```
+
+## Comparing imputation times
+
+TODO:
+1. weź w ramkę danych czasy imputacji po zbiorach (x - id zbioru, y - czas imputacji)
+2. Zrób z tego ggplota z wykresem liniowym (jakiś ładny theme, title, itp. of course)
+3. Wysnuj wnioski o najszybszych i najwolniejszych imputacjach
+
+facet?
+x - id zbioru, y - czas imputacji, z - procent braków
+kategoryczny, ciągły, ciągły
+
+facet?
+x - id zbioru, y - czas imputacji, z - wielkość zbioru
+kategoryczny, ciągły, ciągły
+
+## Comparing modelling times
+
+TODO: to co wyżej, ale na czasach modelowania. Pewnie wniosek, że czas modelowania mega podobny
+
+## Best measures
+
+TODO:
+1. weź w CZTERY ramki danych cztery miary po zbiorach (x - id zbioru, y - wynik miary)
+2. Z pomocą grid_arrange() połącz cztery ggploty z wykresami liniowymi
+3. Wysnuj wnioski o najlepszych i najgorszych miarach na podstawie obrazka
+
+4. Wylicz średnie i wariancje dla wszystkich analizowanych zbiorów dla czterech miar
+5. Zrób prostego ggplota średnich, a imputacji
+6. Zrób prostego ggplota wariancji, a imputacji
+7. Wysnuj wnioski z powyższych dwóch wykresów na podstawie wyliczonych prostych miar
+
+## imputation time and its results
+
+1. Wykres po zbiorach: średnich czasów imputacji a średnich wyników miar.
+2. Wnioski z tego wykresu. Czy są imputacji i lepsze i szybsze? Czy są i wolniejsze i gorsze? PYTANIE: jaki wykres, w jakiej formie?
+
+## Conclusion
+
+TODO: 1. Podsumowanie i wnioski
+
+OPCJONALNIE: porównanie wielkości zbiorów a czasów imputacji

### 2020-04-25 szkielet.r
part res_new

@@ -75,100 +75,65 @@ targets <- lapply(data_all, function(d){d$target})
 
 #PARALLEL ROBI BRRRRR I TEN JAKO JEDYNY DZIAŁA!
 library(parallelMap)
-parallelStartMulticore(8, show.info = TRUE)
+rdzenie <- parallel::detectCores()
+parallelStartMulticore(rdzenie, show.info = TRUE)
 
-res1 <- parallelLapply(data_all, function(x){get_result(x,imputation_remove_rows)}, impute.error = function(x){'ERROR'})
-res2 <- parallelLapply(data_all, function(x){get_result(x,imputation_mode_median)}, impute.error = function(x){'ERROR'})
-res3 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_mice)}, impute.error = function(x){'ERROR'})
-res4 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_missForest)}, impute.error = function(x){'ERROR'})
-res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
+#res1 <- parallelLapply(data_all, function(x){get_result(x,imputation_remove_rows)}, impute.error = function(x){'ERROR'})
+#res2 <- parallelLapply(data_all, function(x){get_result(x,imputation_mode_median)}, impute.error = function(x){'ERROR'})
+#res3 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_mice)}, impute.error = function(x){'ERROR'})
+#res4 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_missForest)}, impute.error = function(x){'ERROR'})
+#res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
 
-
-
-
-### Tu juz nie dziala
-datasets_all <- lapply(data_all, function(x){x$dataset})
-res <- parallelLapply(datasets_all, imputation_fun_mice)
-
-library(parallel)
-
-parallel::setDefaultCluster()
-Job1 = mcparallel(res <- parallelLapply(datasets_all, imputation_fun_mice, impute.error = function(x){'ERROR'} ))
-Job2 = mcparallel(res <- parallelLapply(datasets_all, imputation_remove_rows, impute.error = function(x){'ERROR'}))
[…]

### 2020-04-25 pd2.Rmd

@@
       }
+      else{ # for error input imputation name
+        if (i==3){
+          d[i,1] <- "imputation_fun_mice" 
+        }
+        if (i==4){
+          d[i,1] <- "imputation_fun_missForest" 
+        }
+        if (i==5){
+          d[i,1] <- "imputation_fun_vim" 
+        }
+      }
     }

### 2020-04-25 szkielet.r
update szkielet

@@ -78,31 +78,40 @@ library(parallelMap)
 rdzenie <- parallel::detectCores()
 parallelStartMulticore(rdzenie, show.info = TRUE)
 
+
+#starttime <- Sys.time()
 #res1 <- parallelLapply(data_all, function(x){get_result(x,imputation_remove_rows)}, impute.error = function(x){'ERROR'})
 #res2 <- parallelLapply(data_all, function(x){get_result(x,imputation_mode_median)}, impute.error = function(x){'ERROR'})
 #res3 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_mice)}, impute.error = function(x){'ERROR'})
 #res4 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_missForest)}, impute.error = function(x){'ERROR'})
 #res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
-
-#starttime <- Sys.time()
 #res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
 #endtime <- Sys.time()
-#cat('done in ', endtime - starttime)
+#endtime - starttime
 
 
-saveRDS(res1, file = './part_results/res1_new.rds')
-saveRDS(res2, file = './part_results/res2_new.rds')
-# saveRDS(res3, file = './part_results/res3.rds')
-# saveRDS(res4, file = './part_results/res4.rds')
-# saveRDS(res5, file = './part_results/res5.rds')
+#saveRDS(res1, file = './part_results/res1_new.rds')
+#saveRDS(res2, file = './part_results/res2_new.rds')
+#saveRDS(res3, file = './part_results/res3_new.rds')
+#saveRDS(res4, file = './part_results/res4_new.rds')
+#saveRDS(res5, file = './part_results/res5_new.rds')
 
 
 
-# id 29 naprawione
+# old ids
 v <- c(1590,188,23381,29,38,40536,41278,56,6332)
 
+#id 29
+df29 <- read_dataset(29)
+
+# run kiedy 29 naprawione
+res29 <- parallelLapply(imputations, function(x){get_result(df29, x)}, impute.error = function(x){'ERROR'})
+
+
+# nowe zbiorki
+
 s <- c()
-for(i in 1:14L){
+for(i in 1:length(data_all)){
   s <- c(s, ! (data_all[[i]]$id %in% v))
 }
 
[…]

### 2020-04-26 pd2.Rmd
Jeszcze nie działające nowe zbiory niestety

--- 2020-04-25 pd2.Rmd
@@
 name_of_target <- dataset_list$target
-
 # imputation
@@
 imputation_stop = Sys.time() # end measuring time
-
 # train test split
@@
 test <- as.data.table(train_test[2])
-
 # modelling
@@
 modelling_stop = Sys.time() # end measuring time
-
 # calculating metrics
 confusion_matrix <- get_confusion_matrix(test[[name_of_target]], y_pred[,1])
-
 accuracy_v <- accuracy(confusion_matrix)
@@
 f1_v <- f1(confusion_matrix)
-
 classification_report <- data.frame(accuracy_v, precision_v,
@@
                                      "recall", "f1")
-
 dataset_list$dataset <- NULL
@@
 # for now stick with readability
-
 imp_method_name <- deparse(substitute(imputation_fun))
-
 return(list( dataset_id = dataset_list$id, 
@@
   results <- c(list(0),list(0),list(0),list(0),list(0))
-  results[[1]] <- readRDS('./part_results/res1.rds')
-  results[[2]] <- readRDS('./part_results/res2.rds')
-  results[[3]] <- readRDS('./part_results/res3.rds')
-  results[[4]] <- readRDS('./part_results/res4.rds')
-  results[[5]] <- readRDS('./part_results/res5.rds')
+  results[[1]] <- readRDS('./part_results/res1_new.rds')
+  results[[2]] <- readRDS('./part_results/res2_new.rds')
+  results[[3]] <- readRDS('./part_results/res3_new.rds')
+  results[[4]] <- readRDS('./part_results/res4_new.rds')
+  results[[5]] <- readRDS('./part_results/res5_new.rds')
+
   result_table <- function(ds_it){
@@
   }
+
+  results[[1]]
   # change i bound to number of datasets 
-  for (i in c(1,2,3,5,6,7,8,9)){
+  for (i in c(1, 2, 3, 5, 6, 7, 8, 9)){ # TUTAJ BANGLA DOPÓKI NIE WEŹMIE SIĘ KTÓREGOŚ ZE ZBIORÓW 4, 10, 11, 12, 13, 14 (czyli "nowych")
     kable(result_table(i), caption = paste("Dataset id: ", results[[1]][[i]]$dataset_id, sep = ""))
@@
   kable(result_table(3), caption = paste("Dataset id: ", results[[1]][[3]]$dataset_id, sep = ""))
+  #kable(result_table(4), caption = paste("Dataset id: ", results[[1]][[4]]$dataset_id, sep = ""))
   kable(result_table(5), caption = paste("Dataset id: ", results[[1]][[5]]$dataset_id, sep = ""))
@@
   kable(result_table(9), caption = paste("Dataset id: ", results[[1]][[9]]$dataset_id, sep = ""))
-
+  #kable(result_table(10), caption = paste("Dataset id: ", results[[1]][[10]]$dataset_id, sep = ""))
+  #kable(result_table(11), caption = paste("Dataset id: ", results[[1]][[11]]$dataset_id, sep = ""))
+  #kable(result_table(12), caption = paste("Dataset id: ", results[[1]][[12]]$dataset_id, sep = ""))
+  #kable(result_table(13), caption = paste("Dataset id: ", results[[1]][[13]]$dataset_id, sep = ""))
+  #kable(result_table(14), caption = paste("Dataset id: ", results[[1]][[14]]$dataset_id, sep = ""))
 ```

### 2020-04-27 pd2.Rmd
change getting dataset id from unstable results to median

@@
   for (i in c(1, 2, 3, 5, 6, 7, 8, 9)){ # TUTAJ BANGLA DOPÓKI NIE WEŹMIE SIĘ KTÓREGOŚ ZE ZBIORÓW 4, 10, 11, 12, 13, 14 (czyli "nowych")
-    kable(result_table(i), caption = paste("Dataset id: ", results[[1]][[i]]$dataset_id, sep = ""))
+    kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
   }
@@
 
-  kable(result_table(1), caption = paste("Dataset id: ", results[[1]][[1]]$dataset_id, sep = ""))
-  kable(result_table(2), caption = paste("Dataset id: ", results[[1]][[2]]$dataset_id, sep = ""))
-  kable(result_table(3), caption = paste("Dataset id: ", results[[1]][[3]]$dataset_id, sep = ""))
-  #kable(result_table(4), caption = paste("Dataset id: ", results[[1]][[4]]$dataset_id, sep = ""))
-  kable(result_table(5), caption = paste("Dataset id: ", results[[1]][[5]]$dataset_id, sep = ""))
-  kable(result_table(6), caption = paste("Dataset id: ", results[[1]][[6]]$dataset_id, sep = ""))
-  kable(result_table(7), caption = paste("Dataset id: ", results[[1]][[7]]$dataset_id, sep = ""))
-  kable(result_table(8), caption = paste("Dataset id: ", results[[1]][[8]]$dataset_id, sep = ""))
-  kable(result_table(9), caption = paste("Dataset id: ", results[[1]][[9]]$dataset_id, sep = ""))
-  #kable(result_table(10), caption = paste("Dataset id: ", results[[1]][[10]]$dataset_id, sep = ""))
-  #kable(result_table(11), caption = paste("Dataset id: ", results[[1]][[11]]$dataset_id, sep = ""))
-  #kable(result_table(12), caption = paste("Dataset id: ", results[[1]][[12]]$dataset_id, sep = ""))
-  #kable(result_table(13), caption = paste("Dataset id: ", results[[1]][[13]]$dataset_id, sep = ""))
-  #kable(result_table(14), caption = paste("Dataset id: ", results[[1]][[14]]$dataset_id, sep = ""))
+  kable(result_table(1), caption = paste("Dataset id: ", results[[2]][[1]]$dataset_id, sep = ""))
+  kable(result_table(2), caption = paste("Dataset id: ", results[[2]][[2]]$dataset_id, sep = ""))
+  kable(result_table(3), caption = paste("Dataset id: ", results[[2]][[3]]$dataset_id, sep = ""))
+  #kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
+  kable(result_table(5), caption = paste("Dataset id: ", results[[2]][[5]]$dataset_id, sep = ""))
+  kable(result_table(6), caption = paste("Dataset id: ", results[[2]][[6]]$dataset_id, sep = ""))
+  kable(result_table(7), caption = paste("Dataset id: ", results[[2]][[7]]$dataset_id, sep = ""))
+  kable(result_table(8), caption = paste("Dataset id: ", results[[2]][[8]]$dataset_id, sep = ""))
+  kable(result_table(9), caption = paste("Dataset id: ", results[[2]][[9]]$dataset_id, sep = ""))
+  kable(result_table(10), caption = paste("Dataset id: ", results[[2]][[10]]$dataset_id, sep = ""))
+  kable(result_table(11), caption = paste("Dataset id: ", results[[2]][[11]]$dataset_id, sep = ""))
+  kable(result_table(12), caption = paste("Dataset id: ", results[[2]][[12]]$dataset_id, sep = ""))
+  kable(result_table(13), caption = paste("Dataset id: ", results[[2]][[13]]$dataset_id, sep = ""))
+  kable(result_table(14), caption = paste("Dataset id: ", results[[2]][[14]]$dataset_id, sep = ""))
 ```

### 2020-04-27 pd2.Rmd
4 tez juz bangla

@@
                round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
+               # @TODO konwersja jednostek, as.numeric nie rozroznia jednostek czasu
                round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
@@
   # change i bound to number of datasets 
-  for (i in c(1, 2, 3, 5, 6, 7, 8, 9)){ # TUTAJ BANGLA DOPÓKI NIE WEŹMIE SIĘ KTÓREGOŚ ZE ZBIORÓW 4, 10, 11, 12, 13, 14 (czyli "nowych")
+  for (i in 1:length(results[[1]])){ # TUTAJ BANGLA DOPÓKI NIE WEŹMIE SIĘ KTÓREGOŚ ZE ZBIORÓW 4, 10, 11, 12, 13, 14 (czyli "nowych")
     kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
@@
   kable(result_table(3), caption = paste("Dataset id: ", results[[2]][[3]]$dataset_id, sep = ""))
-  #kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
+  kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
   kable(result_table(5), caption = paste("Dataset id: ", results[[2]][[5]]$dataset_id, sep = ""))

### 2020-04-27 szkielet.r

@@ -66,8 +66,11 @@ get_result <- function(dataset_list, imputation_fun){
 data_all <- read_all_datasets()
 
 # imputations and targets preparation
-imputations <- list(imputation_fun_vim, imputation_fun_missForest,
-                    imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
+imputations <- list(imputation_fun_vim,
+                    imputation_fun_missForest,
+                    imputation_remove_rows,
+                    imputation_mode_median,
+                    imputation_fun_mice)
 
 targets <- lapply(data_all, function(d){d$target})
 
@@ -84,7 +87,6 @@ parallelStartMulticore(rdzenie, show.info = TRUE)
 #res2 <- parallelLapply(data_all, function(x){get_result(x,imputation_mode_median)}, impute.error = function(x){'ERROR'})
 #res3 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_mice)}, impute.error = function(x){'ERROR'})
 #res4 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_missForest)}, impute.error = function(x){'ERROR'})
-#res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
 #res5 <- parallelLapply(data_all, function(x){get_result(x,imputation_fun_vim)}, impute.error = function(x){'ERROR'})
 #endtime <- Sys.time()
 #endtime - starttime

### 2020-04-27 pd2.Rmd
printuje removerows dla pustych ramek

@@
   predM <- init$predictorMatrix
-  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5)
+  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5, nnet.MaxNWts = 2500)
   completed <- complete(imputed)
@@
       else{ # for error input imputation name
+        if (i==1){
+          d[i,1] <- "imputation_remove_rows" 
+        }
+        if (i==2){
+          d[i,1] <- "imputation_mode_median" 
+        }
         if (i==3){

### 2020-04-27 pd2.Rmd
Prawie gotowy wykres (co z legendą?)

--- 2020-04-27 pd2.Rmd
@@
 
-```{r imputation_basic_functions}
+```{r imputation_basic_functions, cache = TRUE}
 imputation_mode_median <- function(df){
@@
 
-```{r imputation_advanced_functions, message = FALSE, warning = FALSE}
+```{r imputation_advanced_functions, message = FALSE, warning = FALSE, cache = TRUE}
 library(mice)
@@
 
-```{r read_dataset}
+```{r read_dataset, cache = TRUE}
 DFT_REPO_DATASET_DIR = './dependencies/datasets'
@@
 
-```{r read_all_datasets}
+```{r read_all_datasets, cache = TRUE}
 read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){
@@
 
-```{r train_test_split}
+```{r train_test_split, cache = TRUE}
 set.seed(1357)
@@
 
-```{r metrics}
+```{r metrics, cache = TRUE}
 get_confusion_matrix <- function(test, pred){
@@
 
-```{r get_result_function, warning = FALSE, message = FALSE}
+```{r get_result_function, warning = FALSE, message = FALSE, cache = TRUE}
 library(rpart)
@@
 
-```{r warning=FALSE}
+```{r warning=FALSE, cache = TRUE}
   library(knitr)
@@
     }
-    d
   }
@@
   # change i bound to number of datasets 
-  for (i in 1:length(results[[1]])){ # TUTAJ BANGLA DOPÓKI NIE WEŹMIE SIĘ KTÓREGOŚ ZE ZBIORÓW 4, 10, 11, 12, 13, 14 (czyli "nowych")
+  for (i in 1:length(results[[1]])){
     kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
@@
 ## Comparing imputation times
+
+```{r imputation_times_counting, cache = TRUE}
+ndatasets <- length(results[[1]])
+nimputations <- length(imputations)
+imputation_names <- c("remove", "median", "mice", "missForest", "VIM")
+dataset_ids <- rep(NULL, ndatasets)
+for(i in 1:ndatasets){
+  dataset_ids[i] <- results[[2]][[i]]$dataset_id
+}
+
+imputation_times <- data.frame(rep(0, 14))
+for (imputation_id in 1:nimputations){
+  imp_id_times <- rep(NULL, ndatasets)
+
+  for(dataset_id in 1:ndatasets){
+    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
+      imp_time <- results[[imputation_id]][[dataset_id]]$imputation_time
+      imp_time <- as.numeric(imp_time, units="secs")
+      imp_id_times[dataset_id] <- imp_time
+    }
+  }
+  imputation_times <- cbind(imputation_times, imp_id_times)
+}
+
+imputation_times[1] <- dataset_ids
+colnames(imputation_times) <- c("dataset_id", imputation_names)
+imputation_times$dataset_id <- as.factor(imputation_times$dataset_id)
+knitr::kable(imputation_times)
+```
+
+```{r imputation_times_plot}
+library(ggplot2)
+
+colors <- c("remove" = "orange", "median" = "cyan", "mice" = "green",
+            "missForest" = "red", "VIM" = "blue")
+
+require(scales)
+ggplot(data = imputation_times, aes(x = dataset_id)) +
+  geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
+  geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
+  geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
+  geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
+  geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
+  ggtitle("Datasets' imputations times") +
+  theme_gray() +
+  scale_y_continuous(trans = log2_trans(),
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +
+  labs(x = "Dataset Id",
+       y = "Imputation time [in seconds]",
+       color = "Legend") +
+  scale_color_manual(name = ":-)", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+```
 

### 2020-04-27 pd2.Rmd
add legend to plot

@@
             "missForest" = "red", "VIM" = "blue")
+imptimes <- melt(imputation_times)
 
 require(scales)
-ggplot(data = imputation_times, aes(x = dataset_id)) +
-  geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
-  geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
-  geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
-  geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
-  geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
+
+# żeby połączyć kolory i shape w legendzie zobacz
+# https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend
+
+ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) + geom_point() + 
+#ggplot(data = imputation_times, aes(x = dataset_id, y )) +
+  #geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
+  #geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
+  #geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
+  #geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
+  #geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
   ggtitle("Datasets' imputations times") +

### 2020-04-27 pd2.Rmd
Comparing imputation times cyyk

--- 2020-04-27 pd2.Rmd
@@
   library(knitr)
-  results <- c(list(0),list(0),list(0),list(0),list(0))
+  results <- rep(list(0), 5)
   results[[1]] <- readRDS('./part_results/res1_new.rds')
@@
     }
+    d
   }
 
-  results[[1]]
   # change i bound to number of datasets 
-  for (i in 1:length(results[[1]])){
-    kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
-  }
-
-  # jest taki problem że kable w pętli nie wypluwa wyniku, chyba że damy print(kable()) wtedy jednak brzydko wygląda
+  #for (i in 1:length(results[[1]])){ # won't use it, because the output looks bad
+  #  kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
+  #}
 
@@
 ## Comparing imputation times
+
+Let's build a dataframe containing information about imputation methods.
 
@@
 
-```{r imputation_times_plot}
+Looks good! Now it's visualization time.
+
+```{r imputation_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
 library(ggplot2)
+# install.packages("reshape") # if not installed
+library(reshape)
+
+# in order to connect colors and shapes in legend see:
+# https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend
 
@@
             "missForest" = "red", "VIM" = "blue")
+imptimes <- melt(imputation_times)
 
 require(scales)
-ggplot(data = imputation_times, aes(x = dataset_id)) +
-  geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
-  geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
-  geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
-  geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
-  geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
+ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) + geom_point() + 
+#ggplot(data = imputation_times, aes(x = dataset_id, y )) +
+  #geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
+  #geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
+  #geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
+  #geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
+  #geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
   ggtitle("Datasets' imputations times") +
@@
        color = "Legend") +
-  scale_color_manual(name = ":-)", values = colors) +
+  scale_color_manual(name = "imputation name", values = colors) +
   theme(legend.position="right",
@@
 
-TODO:
-1. weź w ramkę danych czasy imputacji po zbiorach (x - id zbioru, y - czas imputacji)
-2. Zrób z tego ggplota z wykresem liniowym (jakiś ładny theme, title, itp. of course)
-3. Wysnuj wnioski o najszybszych i najwolniejszych imputacjach
+Judging by the logarithmic scale, no surprise removing rows and median replenishment are definetely the fastest methods, with removing rows being several times faster.
+Looking at more advanced ones, definetely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*.
+These last two are quite slow, with the missforest appearing to be slightly faster.
+
+What do the numbers say?
+
+```{r imputation_times_measures, cache = TRUE}
+means <- rep(NA, nimputations)
+medians <- rep(NA, nimputations)
+stDeviations <- rep(NA, nimputations)
+
+for(i in (1:nimputations)){
+  means[i] <- mean(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
+  means[i] <- round(means[i], 3)
+  medians[i] <- median(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
+  medians[i] <- round(medians[i], 3)
+  stDeviations[i] <- sd(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
+  stDeviations[i] <- round(stDeviations[i], 3)
+}
+
+imputation_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
+knitr::kable(imputation_times_measures)
+```
+
+Taking into acount all the imputations which have been implemented, we can clearly see that two easy ones really stand out.
+However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.
+
+Considering such a large spread of data size, it is very interesting difference between the median and the average for VIM - the first is almost seven minutes, the second - barely one and a half seconds.
+In general, however, it is certainly much longer than methods for removing incomplete rows and filling with median.
+
+Let's count these four measures again, but only for those datasats for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
+
+```{r imputation_times_measures_little, cache = TRUE}
+dataset_ids_fullimp <- c(2, 5, 8, 11, 13, 14)
+ndatasets_fullimp <- length(dataset_ids_fullimp)
+
+means <- rep(NA, nimputations)
+medians <- rep(NA, nimputations)
+stDeviations <- rep(NA, nimputations)
+
+for(i in 1:nimputations){
+  means[i] <- mean(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
+  means[i] <- round(means[i], 3)
+  medians[i] <- median(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
+  medians[i] <- round(medians[i], 3)
+  stDeviations[i] <- sd(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
+  stDeviations[i] <- round(stDeviations[i], 3)
+}
+
+imputation_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
+knitr::kable(imputation_times_measures_fullimp)
+```
+
+Now we have better results to compare.
+Taking into account the mean time, definetely *missForest* is the slowest, but also its standard deviaton is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
+As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.
+
+###########################################3
 

### 2020-04-30 pd2.Rmd
Kubo-czwartkowe poprawki

@@
     code_folding: hide
+  pdf_document: default
 ---
@@
 
-1. *I1* / *modeMedian* - missing data supplemented with medians from individual columns
-2. *I2* / *removeRows* - all the rows with any nonexistent values removed
+1. *modeMedian* - missing data supplemented with medians from individual columns
+2. *removeRows* - all the rows with any nonexistent values removed
 
@@
 
-3. *I3* - *mice*
-4. *I4* - *vim*
-5. *I5* - *missForest*
+3. *mice* - standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
+4. *I4* - *vim* - default *kNN* implementation
+5. *I5* - *missForest* - simple imputation offered by the package
 
@@
 }
+
 ```
@@
 
-* 29 - *credit-approval* PÓKI CO NIE BANGLA
+* 1590 - *adult*
+* 188 - *eucalyptus*
+* 23381 - *dresses-sales*
+* 29 - *credit-approval*
 * 38 - *sick*
-* 56 - ...
-* 188 - *eucalyptus*
-* 1590 - *adult*
-* 6332 - *cylinder-bands*
-* 23381 - *dresses-sales*
 * 40536 - *SpeedDating*
 * 41278 - *okcupid-stem*
+* 56 - *vote*
+* 6332 - *cylinder-bands*
+* 1018 - *ipums_la_99-small*
+* 27 - *colic*
+* 4 - *labor*
+* 55 - *hepatitis*
+* 944 - *echoMonths*
 
@@
 
+These four will give us good information about the quality of imputation and give us a good comparison.
+We get simple and clear information about which observations are well classified, which ones are wrong and how they should be.
+Considering the number of sets to consider and their different properties, even when it comes to balancing, comparing these four measures, the fact that it will show the pros and cons of given imputations, will additionally be good to compare them.
+
 ```{r metrics, cache = TRUE}
@@
 ## Imputation results
+
+The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package.
+We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model.
+The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpast should do a good job.
 
@@
 
-imputation_times <- data.frame(rep(0, 14))
+imputation_times <- data.frame(rep(0, ndatasets))
 for (imputation_id in 1:nimputations){
@@
 require(scales)
-ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) + geom_point() + 
-#ggplot(data = imputation_times, aes(x = dataset_id, y )) +
-  #geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
-  #geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
-  #geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
-  #geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
-  #geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
+ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
+  geom_point() + 
   ggtitle("Datasets' imputations times") +
@@
 
-###########################################3
+## Comparing modeling times
 
-facet?
-x - id zbioru, y - czas imputacji, z - procent braków
-kategoryczny, ciągły, ciągły
+Let's also take a look at modeling times.
 
-facet?
-x - id zbioru, y - czas imputacji, z - wielkość zbioru
-kategoryczny, ciągły, ciągły
+```{r modeling_times_counting, cache = TRUE}
+modeling_times <- data.frame(rep(0, ndatasets))
 
-## Comparing modelling times
+for (imputation_id in 1:nimputations){
+  mod_id_times <- rep(NULL, ndatasets)
 
-TODO: to co wyżej, ale na czasach modelowania. Pewnie wniosek, że czas modelowania mega podobny
+  for(dataset_id in 1:ndatasets){
+    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
+      mod_time <- results[[imputation_id]][[dataset_id]]$modelling_time
+      mod_time <- as.numeric(mod_time, units="secs")
+      mod_id_times[dataset_id] <- mod_time
+    }
+  }
+  modeling_times <- cbind(modeling_times, mod_id_times)
+}
+
+modeling_times[1] <- dataset_ids
+colnames(modeling_times) <- c("dataset_id", imputation_names)
+imputation_times$dataset_id <- as.factor(modeling_times$dataset_id)
+knitr::kable(modeling_times)
+```
+
+Very interesting!
+The longest modeling time took one of the largest sets...
+on deficiencies supplemented by median.
+And it took apparently longer than other modeling!
+could a random tree have more to do with a lot of the same values?
+Probably yes.
+Let's visualize our results.
+
+```{r modeling_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
+modeling_times$dataset_id <- as.factor(modeling_times$dataset_id)
+modtimes <- melt(modeling_times)
+
+require(scales)
+ggplot(data=modtimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
+  geom_point() + 
+  ggtitle("Datasets' modeling times") +
+  theme_light() +
+  scale_y_continuous(trans = log2_trans(),
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +
+  labs(x = "Dataset Id",
+       y = "Modeling time [in seconds]",
+       color = "Legend") +
+  scale_color_manual(name = "imputation name", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+```
+
+Here the results seem to be way more scattered. Let's check on measures also.
+
+```{r modeling_times_measures, cache = TRUE}
+means <- rep(NA, nimputations)
+medians <- rep(NA, nimputations)
+stDeviations <- rep(NA, nimputations)
+
+for(i in (1:nimputations)){
+  means[i] <- mean(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
+  means[i] <- round(means[i], 3)
+  medians[i] <- median(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
+  medians[i] <- round(medians[i], 3)
+  stDeviations[i] <- sd(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
+  stDeviations[i] <- round(stDeviations[i], 3)
+}
+
+modeling_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
+knitr::kable(modeling_times_measures)
+```
+
+The powerful but also fast tree classifier from *rpast* (both looking at the median and average) counted everything in less than a second.
+The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds.
+You can also see very similar all three average measures for these two, probably due to the operation of *VIM* as a result giving similar effects as the *median*.
+Interestingly, unbeatably the fastest tree was built for the *mice* algorithm ...
+but wait!
+Maybe we'd better compare only the imputations for those sets for which we've done everything.
+
+```{r modeling_times_measures_little, cache = TRUE}
+means <- rep(NA, nimputations)
+medians <- rep(NA, nimputations)
+stDeviations <- rep(NA, nimputations)
+
+for(i in 1:nimputations){
+  means[i] <- mean(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
+  means[i] <- round(means[i], 3)
+  medians[i] <- median(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
+  medians[i] <- round(medians[i], 3)
+  stDeviations[i] <- sd(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
+  stDeviations[i] <- round(stDeviations[i], 3)
+}
+
+modeling_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
+knitr::kable(modeling_times_measures_fullimp)
+```
+
+On the attached non-manipulated table, all average measures are almost identical.
+The conclusion is that the method of imputation probably has almost no effect on the modeling time - although it is difficult to say, given that we had a fairly small number of sets to compare and we have values in the order of hundredths of a second - probably just the operation of the computer pretty badly distorts the result.
 
@@
 
-TODO:
-1. weź w CZTERY ramki danych cztery miary po zbiorach (x - id zbioru, y - wynik miary)
-2. Z pomocą grid_arrange() połącz cztery ggploty z wykresami liniowymi
-3. Wysnuj wnioski o najlepszych i najgorszych miarach na podstawie obrazka
+Let us draw attention to our measures. How did our four measures work on them?
 
-4. Wylicz średnie i wariancje dla wszystkich analizowanych zbiorów dla czterech miar
-5. Zrób prostego ggplota średnich, a imputacji
-6. Zrób prostego ggplota wariancji, a imputacji
-7. Wysnuj wnioski z powyższych dwóch wykresów na podstawie wyliczonych prostych miar
+Attention!
+We will not consider sets with id 41278 and 188 in the next steps - they are non-binary classifications, so they are not affected by accuracy, precision, recall and f1.
+However, from the remaining twelve collections - we will be able to draw many valuable conclusions.
 
-## imputation time and its results
+```{r measures_tables, cache = TRUE}
+bin_ids <- c(1, 3, 4, 5, 6, 8, 9, 10, 11, 12, 13, 14)
 
-1. Wykres po zbiorach: średnich czasów imputacji a średnich wyników miar.
-2. Wnioski z tego wykresu. Czy są imputacji i lepsze i szybsze? Czy są i wolniejsze i gorsze? PYTANIE: jaki wykres, w jakiej formie?
+accuracies <- data.frame(rep(0, ndatasets))
+precisions <- data.frame(rep(0, ndatasets))
+recalls <- data.frame(rep(0, ndatasets))
+f1s <- data.frame(rep(0, ndatasets))
+
+for (imputation_id in 1:nimputations){
+  acc_id <- rep(NA, ndatasets)
+  pre_id <- rep(NA, ndatasets)
+  rec_id <- rep(NA, ndatasets)
+  f1_id <- rep(NA, ndatasets)
+
+  for(dataset_id in bin_ids){
+    if("classification_report" %in% names(results[[imputation_id]][[dataset_id]])){
+      accuracy <- results[[imputation_id]][[dataset_id]]$classification_report$accuracy
+      acc_id[dataset_id] <- accuracy
+      precision <- results[[imputation_id]][[dataset_id]]$classification_report$precision
+      pre_id[dataset_id] <- precision
+      recall <- results[[imputation_id]][[dataset_id]]$classification_report$recall
+      rec_id[dataset_id] <- recall
+      f1 <- results[[imputation_id]][[dataset_id]]$classification_report$f1
+      f1_id[dataset_id] <- f1
+    }
+  }
+  accuracies <- cbind(accuracies, acc_id)
+  precisions <- cbind(precisions, pre_id)
+  recalls <- cbind(recalls, rec_id)
+  f1s <- cbind(f1s, f1_id)
+}
+
+accuracies[1] <- dataset_ids
+colnames(accuracies) <- c("dataset_id", imputation_names)
+accuracies$dataset_id <- as.factor(accuracies$dataset_id)
+accuracies <- accuracies[-c(2, 7), ] # delete  non-binary classification datasets
+knitr::kable(accuracies, caption = "Accuracies")
+
+precisions[1] <- dataset_ids
+colnames(precisions) <- c("dataset_id", imputation_names)
+precisions$dataset_id <- as.factor(precisions$dataset_id)
+precisions <- precisions[-c(2, 7), ] # delete  non-binary classification datasets
+knitr::kable(precisions, caption = "Precisions")
+
+recalls[1] <- dataset_ids
+colnames(recalls) <- c("dataset_id", imputation_names)
+recalls$dataset_id <- as.factor(recalls$dataset_id)
+recalls <- recalls[-c(2, 7), ] # delete  non-binary classification datasets
+knitr::kable(recalls, caption = "Recalls")
+
+f1s[1] <- dataset_ids
+colnames(f1s) <- c("dataset_id", imputation_names)
+f1s$dataset_id <- as.factor(f1s$dataset_id)
+f1s <- f1s[-c(2, 7), ] # delete  non-binary classification datasets
+knitr::kable(f1s, caption = "F1s")
+```
+
+Much data!!
+
+Naturally - we'll visualize it.
+
+```{r measures_visualizations, message = FALSE, warning = FALSE, cache = TRUE}
+accMelt <- melt(accuracies)
+accPlot <- ggplot(data=accMelt, aes(x = dataset_id, y = value, color = variable)) +
+  geom_point() + 
+  ggtitle("Datasets' accuracies") +
+  ylim(0, 1) +
+  labs(x = "Dataset Id",
+       y = "Accuracy",
+       color = "Legend") +
+  scale_color_manual(name = "accuracy", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+accPlot
+
+preMelt <- melt(precisions)
+prePlot <- ggplot(data=preMelt, aes(x = dataset_id, y = value, color = variable)) +
+  geom_point() + 
+  ggtitle("Datasets' precisions") +
+  ylim(0, 1) +
+  labs(x = "Dataset Id",
+       y = "Precision",
+       color = "Legend") +
+  scale_color_manual(name = "precision", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+prePlot
+
+recMelt <- melt(recalls)
+recPlot <- ggplot(data=recMelt, aes(x = dataset_id, y = value, color = variable)) +
+  geom_point() + 
+  ggtitle("Datasets' recalls") +
+  ylim(0, 1) +
+  labs(x = "Dataset Id",
+       y = "Recall",
+       color = "Legend") +
+  scale_color_manual(name = "recall", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+recPlot
+
+f1Melt <- melt(f1s)
+f1Plot <- ggplot(data=f1Melt, aes(x = dataset_id, y = value, color = variable)) +
+  geom_point() + 
+  ggtitle("Datasets' f1s") +
+  ylim(0, 1) +
+  labs(x = "Dataset Id",
+       y = "F1",
+       color = "Legend") +
+  scale_color_manual(name = "f1", values = colors) +
+  theme(legend.position="right",
+        plot.title = element_text(hjust = 0.5))
+f1Plot
+
+library(gridExtra)
+grid.arrange(accPlot +
+               ggtitle("Accuracies") +
+               theme(legend.position = "none",
+                axis.line.x = element_blank(),
+                axis.ticks.x = element_blank(),
+                axis.text.x = element_blank(),
+                axis.title = element_blank()),
+              prePlot +
+               ggtitle("Precisions") +
+               theme(legend.position = "none",
+                axis.line.x = element_blank(),
+                axis.ticks.x = element_blank(),
+                axis.text.x = element_blank(),
+                axis.title = element_blank()),
+              recPlot +
+               ggtitle("Recalls") +
+               theme(legend.position = "none",
+                axis.line.x = element_blank(),
+                axis.ticks.x = element_blank(),
+                axis.text.x = element_blank(),
+                axis.title = element_blank()),
+             f1Plot +
+               ggtitle("F1s") +
+               theme(legend.position = "none",
+                axis.line.x = element_blank(),
+                axis.ticks.x = element_blank(),
+                axis.text.x = element_blank(),
+                axis.title = element_blank()),
+             nrow = 2,
+             top = "Measues values for different datasets' imputations")
+```
+
+... where colors of points represent: <p style="color:orange">*1. **remove***</p>
+<p style="color:cyan">*2. **median***</p><p style="color:green">*3. **mice***</p>
+<p style="color:red">*4. **missForest***</p><p style="color:blue">*5. **VIM***</p>
+
+Looking only at these plots, it is difficult to come up with something at first glance - for each time I come to some conclusion, results for a few other datasets that I did not pay attention to deny it.
+For sure we can already say that the scatter of results is rather large and analyzing the measures of concentration, we probably will not clearly indicate the best or the worst imputation.
+
+Visualizations often help and ever give answers to the questions, but this time the best option will be definetely to calculate some measures.
+What means, medians and standard deviations give these measures?
+
+```{r measures_measures, cache = TRUE}
+accMeans <- rep(NA, nimputations)
+accMedians <- rep(NA, nimputations)
+accStDeviations <- rep(NA, nimputations)
+
+preMeans <- rep(NA, nimputations)
+preMedians <- rep(NA, nimputations)
+preStDeviations <- rep(NA, nimputations)
+
+recMeans <- rep(NA, nimputations)
+recMedians <- rep(NA, nimputations)
+recStDeviations <- rep(NA, nimputations)
+
+f1sMeans <- rep(NA, nimputations)
+f1sMedians <- rep(NA, nimputations)
+f1sStDeviations <- rep(NA, nimputations)
+
+for(i in (1:nimputations)){
+    accMeans[i] <- mean(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
+    accMeans[i] <- round(accMeans[i], 3)
+    accMedians[i] <- median(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
+    accMedians[i] <- round(accMedians[i], 3)
+    accStDeviations[i] <- sd(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
+    accStDeviations[i] <- round(stDeviations[i], 3)
+
+    preMeans[i] <- mean(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
+    preMeans[i] <- round(preMeans[i], 3)
+    preMedians[i] <- median(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
+    preMedians[i] <- round(preMedians[i], 3)
+    preStDeviations[i] <- sd(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
+    preStDeviations[i] <- round(preStDeviations[i], 3)
+
+    recMeans[i] <- mean(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
+    recMeans[i] <- round(recMeans[i], 3)
+    recMedians[i] <- median(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
+    recMedians[i] <- round(recMedians[i], 3)
+    recStDeviations[i] <- sd(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
+    recStDeviations[i] <- round(recStDeviations[i], 3)
+
+    f1sMeans[i] <- mean(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
+    f1sMeans[i] <- round(f1sMeans[i], 3)
+    f1sMedians[i] <- median(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
+    f1sMedians[i] <- round(f1sMedians[i], 3)
+    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
+    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
+}
+
+measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
+                                preMeans, preMedians, preStDeviations,
+                                recMeans, recMedians, recStDeviations,
+                                f1sMeans, f1sMedians, f1sStDeviations)
+
+knitr::kable(measures_measures, content = "Model measures' concentrations' measures")
+```
+
+Looking at the table quickly, we can probably set our podium!!
+*MissForest* becomes the big winner just before ex aequo in second place: *mice* and *VIM*.
+At the bottom we have *median*, slightly overtaken by *remove_rows*?
+Well no!
+Remember that we should not take datasets for such comparison, for which we have only partial results.
+They probably strongly disturb the information about the data.
+The table above is not useless, because in the end we have means, medians and standard deviations of everything we could; but in order to be able to compare them without remorse, we should take only these five - which have complete information - we will calculate again the measures for datasets with ids 38, 56, 27, 55 and 944.
+This is necessary, because if not for the sets in which, naturally, regardless of the model or imputation we get very high or very low results generate a lot of chaos.
+
+```{r correct_measures_measures, cache = TRUE}
+ids_all_model_measures <- c(4, 6, 9, 11, 12)
+
+accMeans <- rep(NA, nimputations)
+accMedians <- rep(NA, nimputations)
+accStDeviations <- rep(NA, nimputations)
+
+preMeans <- rep(NA, nimputations)
+preMedians <- rep(NA, nimputations)
+preStDeviations <- rep(NA, nimputations)
+
+recMeans <- rep(NA, nimputations)
+recMedians <- rep(NA, nimputations)
+recStDeviations <- rep(NA, nimputations)
+
+f1sMeans <- rep(NA, nimputations)
+f1sMedians <- rep(NA, nimputations)
+f1sStDeviations <- rep(NA, nimputations)
+
+for(i in 1:nimputations){
+    accMeans[i] <- mean(accuracies[imputation_names[i]][ids_all_model_measures,])
+    accMeans[i] <- round(accMeans[i], 3)
+    accMedians[i] <- median(accuracies[imputation_names[i]][ids_all_model_measures,])
+    accMedians[i] <- round(accMedians[i], 3)
+    accStDeviations[i] <- sd(accuracies[imputation_names[i]][ids_all_model_measures,])
+    accStDeviations[i] <- round(stDeviations[i], 3)
+
+    preMeans[i] <- mean(precisions[imputation_names[i]][ids_all_model_measures,])
+    preMeans[i] <- round(preMeans[i], 3)
+    preMedians[i] <- median(precisions[imputation_names[i]][ids_all_model_measures,])
+    preMedians[i] <- round(preMedians[i], 3)
+    preStDeviations[i] <- sd(precisions[imputation_names[i]][ids_all_model_measures,])
+    preStDeviations[i] <- round(preStDeviations[i], 3)
+
+    recMeans[i] <- mean(recalls[imputation_names[i]][ids_all_model_measures,])
+    recMeans[i] <- round(recMeans[i], 3)
+    recMedians[i] <- median(recalls[imputation_names[i]][ids_all_model_measures,])
+    recMedians[i] <- round(recMedians[i], 3)
+    recStDeviations[i] <- sd(recalls[imputation_names[i]][ids_all_model_measures,])
+    recStDeviations[i] <- round(recStDeviations[i], 3)
+
+    f1sMeans[i] <- mean(f1s[imputation_names[i]][ids_all_model_measures,])
+    f1sMeans[i] <- round(f1sMeans[i], 3)
+    f1sMedians[i] <- median(f1s[imputation_names[i]][ids_all_model_measures,])
+    f1sMedians[i] <- round(f1sMedians[i], 3)
+    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][ids_all_model_measures,])
+    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
+}
+
+correct_measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
+                                preMeans, preMedians, preStDeviations,
+                                recMeans, recMedians, recStDeviations,
+                                f1sMeans, f1sMedians, f1sStDeviations)
+
+knitr::kable(correct_measures_measures, content = "CORRECT model measures' concentrations' measures")
+```
+
+So now we can make some conclusions!
+
+Having the table above, we can't say that *missForest* is definitely the best.
+Looking at the advanced ones, *missForest* and *VIM* are doing very well, almost always better than their competitors.
+*VIM* has slightly better accuracy and definitely better precision, while *missForest* dominates in the recall case and has slightly better f1.
+***Mice* is almost always the worst**.
+Even *remove_rows* is doing better than it, and the *median* sometimes catches up with missForest and mice.
+
+Standard deviations are pretty balanced and it's hard to tell anything clever about it - especially considering the small number of collections. Rather, nothing stands out.
+
+Let's visualize dataset's model result's measures.
+
+```{r model_measures_plots, message = FALSE, cache = TRUE}
+colors2 <- c("mean" = "blue", "median" = "orange")
+
+acc_measures <- correct_measures_measures %>% select(imputation_names, accMeans, accMedians)
+colnames(acc_measures) <- c("imputation_name", "mean", "median")
+accMelt <- melt(acc_measures)
+accuracyPlot <- ggplot(data = accMelt, aes(x = imputation_name, y = value, color = variable)) +
+  geom_point(size = 3) + 
+  ylim(0.75, 1) +
+  theme_light() +
+  ggtitle("Accuracies") +
+  labs(color = "Legend") +
+  scale_color_manual(name = "measure", values = colors2) +
+  theme(legend.position="None",
+        plot.title = element_text(hjust = 0.5),
+        axis.title = element_blank())
+
+pre_measures <- correct_measures_measures %>% select(imputation_names, preMeans, preMedians)
+colnames(pre_measures) <- c("imputation_name", "mean", "median")
+preMelt <- melt(pre_measures)
+precisionPlot <- ggplot(data = preMelt, aes(x = imputation_name, y = value, color = variable)) +
+  geom_point(size = 3) + 
+  ylim(0.75, 1) +
+  theme_light() +
+  ggtitle("Precisions") +
+  labs(color = "Legend") +
+  scale_color_manual(name = "measure", values = colors2) +
+  theme(legend.position="None",
+        plot.title = element_text(hjust = 0.5),
+        axis.title = element_blank())
+
+rec_measures <- correct_measures_measures %>% select(imputation_names, preMeans, preMedians)
+colnames(rec_measures) <- c("imputation_name", "mean", "median")
+recMelt <- melt(rec_measures)
+recallPlot <- ggplot(data = recMelt, aes(x = imputation_name, y = value, color = variable)) +
+  geom_point(size = 3) + 
+  ylim(0.75, 1) +
+  theme_light() +
+  ggtitle("Recalls") +
+  labs(x = "Imputation name",
+       y = "Recall",
+       color = "Legend") +
+  scale_color_manual(name = "measure", values = colors2) +
+  theme(legend.position="None",
+        plot.title = element_text(hjust = 0.5),
+        axis.title = element_blank())
+
+f1_measures <- correct_measures_measures %>% select(imputation_names, f1sMeans, f1sMedians)
+colnames(f1_measures) <- c("imputation_name", "mean", "median")
+f1mMelt <- melt(f1_measures)
+f1sPlot <- ggplot(data = f1mMelt, aes(x = imputation_name, y = value, color = variable)) +
+  geom_point(size = 3) + 
+  ylim(0.75, 1) +
+  theme_light() +
+  ggtitle("F1s") +
+  labs(x = "Imputation name",
+       y = "f1",
+       color = "Legend") +
+  scale_color_manual(name = "measure", values = colors2) +
+  theme(legend.position="None",
+        plot.title = element_text(hjust = 0.5),
+        axis.title = element_blank())
+
+grid.arrange(accuracyPlot, precisionPlot,
+             recallPlot, f1sPlot, nrow = 2,
+             top = "Means (blue) and medians (orange) of datasets' imputations' modeling results")
+```
+
+I think we will all agree now - in terms of results, at least looking at these six datasets for which each imputation was successful, we can create such a podium:
+
+1. *VIM* / *missForest* (judging by our need - for *accuracy* and *precision* - *VIM*, for *recall* - *missForest*)
+3. *median*
+4. *remove_rows*
+5. *mice*
+
+Of course, the results are not very representative, given the small number of datasets for which they were compared - but still something.
+
+## Imputation time and its results
+
+We collected and summarized both results and imputation times in our collections. What are the average imputation times and the measures' medians of our imputations?
+
+```{r imputation_times_measures_medians, cache = TRUE}
+medians_measures_times <- correct_measures_measures %>% select(imputation_names, accMedians, preMedians, recMedians, f1sMedians)
+medians_measures_times <- cbind(medians_measures_times, imputation_times_measures_fullimp$medians)
+colnames(medians_measures_times) <- c("imputation name", "accuracy", "precision", "recall", "f1", "imputation mean time")
+knitr::kable(medians_measures_times, content = "Mean imputation time and mean model result's measures of imp methods")
+```
+
+Looking at the results, *VIM* and *missForest* give quite similar numbers in terms of satisfaction, but VIM is much faster - for data sets for which each imputation has passed, it managed in six times shorter period of time than *missForest* (and also of course much worse *mice*).
+The question is how much we are interested in time - it turns out that for very large data sets our model will not suffer much if we take the median instead of these two.
+At this point, you can probably make a correction of the ranking:
+
+1. *VIM*
+2. *median* / *missForest* (depending on how important time is for us)
+...
+4. *remove_rows*
+...
+5. *mice*
 
@@
 
-TODO: 1. Podsumowanie i wnioski
-
-OPCJONALNIE: porównanie wielkości zbiorów a czasów imputacji
+The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a very good package, supplementing the missing values with medians from columns is not at all is such a stupid idea as it may seem, and *mice* can be thrown in the bin.
+Unfortunately, it is impossible to draw a wise conclusion based on such a small number of tested data - admittedly as many as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer and assumptions that respectful statistics would praise.
+Nevertheless, above code is certainly very valuable and for more data sets it could confirm one or another belief - therefore we can probably say that we have achieved scientific success!

### 2020-04-30 pd2.Rmd
wstępne poprawki

@@
 
-The eight data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research.
+The fourteen data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research.
 They can be found under the following identifiers with the following names:
@@
 
-In order to read all the eight datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.
+In order to read all datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.
 
@@
 }
+```
+
+All of datasets contain missing information denoted as NA and differ in size, number of features, number of observations and percentage of missing data. Table below decribes those informations.
+
+```{r}
+### TU KOD HANI BEDZIE DO WKLEJENIA
+
+dfs <- read_all_datasets()
 ```
@@
 We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model.
-The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpast should do a good job.
+The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.
 
@@
 
-Judging by the logarithmic scale, no surprise removing rows and median replenishment are definetely the fastest methods, with removing rows being several times faster.
-Looking at more advanced ones, definetely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*.
+Judging by the logarithmic scale, no surprise removing rows and median replenishment are definitely the fastest methods, with removing rows being several times faster.
+Looking at more advanced ones, definitely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*.
 These last two are quite slow, with the missforest appearing to be slightly faster.
@@
 
-Taking into acount all the imputations which have been implemented, we can clearly see that two easy ones really stand out.
+Taking into account all the imputations which have been implemented, we can clearly see that two easy ones really stand out.
 However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.
@@
 
-Let's count these four measures again, but only for those datasats for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
+Let's count these four measures again, but only for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
 
@@
 Now we have better results to compare.
-Taking into account the mean time, definetely *missForest* is the slowest, but also its standard deviaton is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
+Taking into account the mean time, definitely *missForest* is the slowest, but also its standard deviation is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
 As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.
@@
 
-The powerful but also fast tree classifier from *rpast* (both looking at the median and average) counted everything in less than a second.
+The powerful but also fast tree classifier from *rpart* (both looking at the median and average) counted everything in less than a second.
 The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds.
@@
              nrow = 2,
-             top = "Measues values for different datasets' imputations")
+             top = "Measures values for different datasets' imputations")
 ```
@@
 
-Visualizations often help and ever give answers to the questions, but this time the best option will be definetely to calculate some measures.
+Visualizations often help and ever give answers to the questions, but this time the best option will be definitely to calculate some measures.
 What means, medians and standard deviations give these measures?
@@
 
-Let's visualize dataset's model result's measures.
+Let's visualize dataset's model results measures.
 

### 2020-04-30 pd2.Rmd
added tabelka z opisem zbiorow

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
 ```
@@
 
-```{r}
+```{r dataset description, message = FALSE, warning=FALSE, cache = TRUE}
 ### TU KOD HANI BEDZIE DO WKLEJENIA
@@
 dfs <- read_all_datasets()
+
+ids <- lapply(dfs, function(x){x$id})
+names <- lapply(dfs, function(x){x$data_name})
+n_inst <- lapply(dfs, function(x){x$number_of_instances})
+n_features <- lapply(dfs, function(x){x$number_of_features})
+number_of_numeric_features <- lapply(dfs, function(x){x$number_of_numeric_features})
+number_of_categorical_features <- lapply(dfs, function(x){x$number_of_categorical_features})
+number_of_missings <- lapply(dfs, function(x){x$number_of_missings})
+proc_missing <- lapply(dfs, function(x){x$number_of_missing/(x$number_of_instances * x$number_of_features)})
+number_of_instances_with_missings <- lapply(dfs, function(x){x$number_of_instances_with_missings})
+number_of_features_with_missings <- lapply(dfs, function(x){x$number_of_features_with_missings})
+
+#fix wrong json
+names[[10]] <- "okcupid-stem"
+
+table <- data.frame("id" = unlist(ids), 
+                    "name" = unlist(names),
+                    "number_of_instances" = unlist(n_inst),
+                    "number_of_features" = unlist(n_features),
+                    "number_of_missing" = unlist(number_of_missings),
+                    "procentage_of_missing" = round(unlist(proc_missing)*100,2))
+
+knitr::kable(table)
 ```

### 2020-04-30 pd2.Rmd
poprawiona legenda

@@
 ```{r dataset description, message = FALSE, warning=FALSE, cache = TRUE}
-### TU KOD HANI BEDZIE DO WKLEJENIA
 
@@
 
-Here are the results:
+Here are the results. NA indicates that method returned error:
 
@@
   theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
+        plot.title = element_text(hjust = 0.5)) +
+  #scale_colour_manual(name = "Treatment & State",
+  #                    labels = c("Control, Non-F", "Control, Flwr", "Exclosure, Non-F", "Exclosure, Flwr"),
+  #                    values = c("blue", "red", "blue", "red")) +   
+  scale_shape_manual(name = "imputation name",
+                     values = 15:19L)
+
 ```

### 2020-05-20 article.Rmd
Add files via upload

--- 2020-04-30 pd2.Rmd
@@
 ---
-title: "The imputation methods big test"
+title: "Jaki tytuł??"
 author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
-date: "22/04/2020"
+date: "20/05/2020"
 output:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
 ```
 
-## Abstract
+# Abstract
 
 Choosing the best-suited imputation is the daily dilemma of every data scientist.
-Some of they believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.
+Some of them believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.
 
-In this document we will try to find one objectively the finest imputation method by testing five of popular ones on eight specially selected data sets prepared for these operations.
+In this document we will try to find the one, objectively finest imputation method by testing six of popular ones on forteen specially selected data sets prepared for these operations.
+
+# Introduction + motivation
 
@@
 
-We'll look at the two simple imputation methods with the following short markings:
+We'll look at the three simple imputation methods and three more advanced algorithms offered by popular packages.
 
-1. *modeMedian* - missing data supplemented with medians from individual columns
-2. *removeRows* - all the rows with any nonexistent values removed
+TODO: czemu one; lanie wody o tym, że są fajne bo różne i badamy różne spojrzenia
 
-```{r imputation_basic_functions, cache = TRUE}
-imputation_mode_median <- function(df){
+1. *mode / median replace* - basic method which puts in a place of missing data median cells from their columns for numerical values and dominants for categorical ones.
+2. *remove rows* - trivial solution of removing rows containing any missing data.
+3. *bogo replace* - simple algorithm replacing NA values with random numbers or factors from their columns.
+4.
+*mice* ^1^ - method creating multiple imputations for multivariate missing data, based on Fully Conditional Specification, where each incomplete variable is imputed by a separate model.
+Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
+5. *VIM* ^2^ - standard k-nearest neighbors algorithm taken from the library
+6. *missForest* ^3^ - simple imputation based on random forest offered by the package
 
-  Mode <- function(x) {
-    ux <- unique(x)
-    ux[which.max(tabulate(match(x, ux)))]
-  }
+## Datasets
 
-  for (i in 1L:length(df)){
-    if (sum(is.na(df[,i])) > 0){
-      if (mode(df[,i]) == 'character' | is.factor(df[,i])){
-        to_imp <- Mode(df[,i])
-        df[,i][is.na(df[,i])] <- to_imp
-      }
-      else{
-        to_imp <- median(df[,i], na.rm = TRUE) 
-        df[,i][is.na(df[,i])] <- to_imp
-      }
-    }
-  }
-
-  return(df)
-}
-imputation_remove_rows <- function(df){
-  return (na.omit(df))
-}
-```
-
-... and also three more advanced algorithms from popular libraries:
-
-3. *mice* - standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
-4. *I4* - *vim* - default *kNN* implementation
-5. *I5* - *missForest* - simple imputation offered by the package
-
-```{r imputation_advanced_functions, message = FALSE, warning = FALSE, cache = TRUE}
-library(mice)
-imputation_fun_mice <- function(df){
-  init <- mice(df, maxit=0) 
-  meth <- init$method
-  predM <- init$predictorMatrix
-  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5)
-  completed <- complete(imputed)
-  return(completed)
-}
-library(VIM)
-imputation_fun_vim <- function(df){
-  no_columns <- length(df)
-  imputed <- kNN(df)
-  imputed <- imputed[,1:no_columns]
-  return(imputed)
-}
-library(missForest)
-imputation_fun_missForest <- function(df){
-  return(missForest(df)$ximp)
-}
-
-```
-
-## Reading datasets
+TODO: ładniejszy format
 
@@
 
-```{r read_dataset, cache = TRUE}
-DFT_REPO_DATASET_DIR = './dependencies/datasets'
-read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){
+TODO: tylko na sześciu wszystko się udało
 
-  if (!dir.exists(dataset_dir)){
-    stop(paste(dataset_dir, 'does not exist' ))
-  }
+## Binary classification algorithms
 
-  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
-  if (!dir.exists(dir)){
-    stop(paste(dir, 'does not exist' ))
-  }
+TODO: pięć metod, ich krótki opis i również krótkie stwierdzenie, że sa fajne bo różne i wgl.
 
-  start_dir <- getwd()
+# Description of methodology and results
 
-  # set right dir to code.R to acually work - it depends on dirlocation to create json
-  setwd(dir)
-  # use new env to avoid trashing globalenv
-  surogate_env <- new.env(parent = .BaseNamespaceEnv)
-  attach(surogate_env)
-  source("code.R",surogate_env)
+## Reading datasets
 
-  j <- jsonlite::read_json('./dataset.json')
-  j$dataset <- surogate_env$dataset
-  setwd(start_dir)
+In order to read all datasets we used another function, which returned a dataframe containing all the important informations about the test matrices.
 
-  return(j)
-}
-```
-
-In order to read all datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.
-
-```{r read_all_datasets, cache = TRUE}
-read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){
-
-  if (!dir.exists(dataset_dir)){
-    stop(paste(dataset_dir, 'does not exist'))
-  }
-
-  start_dir <- getwd()
-  subdirs <- dir(dataset_dir)
-  ids <- sapply(subdirs, function(dir){substr(dir, 16, nchar(dir))})
-  datasets_combined <- lapply(ids, function(x){read_dataset(x, dataset_dir)})
-
-  datasets_combined <- t(datasets_combined)
-  return(unname(t(datasets_combined)))
-}
-```
-
-All of datasets contain missing information denoted as NA and differ in size, number of features, number of observations and percentage of missing data. Table below decribes those informations.
-
-```{r}
-### TU KOD HANI BEDZIE DO WKLEJENIA
-
-dfs <- read_all_datasets()
-```
+TODO: trochę o metodzie, o funkcjach czytających; PARALLEL, itd
 
@@
 In order to test the same test-train splits we'll use one random seed 1357 for all datasets.
-
-```{r train_test_split, cache = TRUE}
-set.seed(1357)
-train_test_split <- function(dataset, train_size){
-  smp_size <- floor(train_size * nrow(dataset))
-  typeof(smp_size)
-
-  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)
-
-  train <- dataset[train_ind, ]
-  test <- dataset[-train_ind, ]
-
-  return (list(train, test))
-}
-```
 
@@
 
-```{r metrics, cache = TRUE}
-get_confusion_matrix <- function(test, pred){
-  return (table(Truth = test, Prediction = pred))
-}
-confusion_matrix_values <- function(confusion_matrix){
-  TP <- confusion_matrix[2,2]
-  TN <- confusion_matrix[1,1]
-  FP <- confusion_matrix[1,2]
-  FN <- confusion_matrix[2,1]
-  return (c(TP, TN, FP, FN))
-}
-accuracy <- function(confusion_matrix){
-  conf_matrix <- confusion_matrix_values(confusion_matrix)
-  return((conf_matrix[1] + conf_matrix[2]) / (conf_matrix[1] + conf_matrix[2] + conf_matrix[3] + conf_matrix[4]))
-}
-precision <- function(confusion_matrix){
-  conf_matrix <- confusion_matrix_values(confusion_matrix)
-  return(conf_matrix[1]/ (conf_matrix[1] + conf_matrix[3]))
-}
-recall <- function(confusion_matrix){
-  conf_matrix <- confusion_matrix_values(confusion_matrix)
-  return(conf_matrix[1] / (conf_matrix[1] + conf_matrix[4]))
-}
-f1 <- function(confusion_matrix){
-  conf_matrix <- confusion_matrix_values(confusion_matrix)
-  rec <- recall(confusion_matrix)
-  prec <- precision(confusion_matrix)
-  return(2 * (rec * prec) / (rec + prec))
-}
-```
+TODO: MCC jeszcze, czemu f1
 
@@
 
-The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package.
-We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model.
+The results of our work will be checked using the implemented classification tree algorithm from the *rpart ^4^* package.
+We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a decent model.
 The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.
 
-```{r get_result_function, warning = FALSE, message = FALSE, cache = TRUE}
-library(rpart)
-get_result <- function(dataset_list, imputation_fun){
-
-dataset <- dataset_list$dataset
-name_of_target <- dataset_list$target
-# imputation
-imputation_start = Sys.time() # start to measure time
-imputated_dataset <- imputation_fun(dataset) 
-imputation_stop = Sys.time() # end measuring time
-# train test split
-train_test <- train_test_split(imputated_dataset, 0.8)
-train <- as.data.table(train_test[1])
-test <- as.data.table(train_test[2])
-# modelling
-vars <- colnames(dataset)[colnames(dataset)!=name_of_target]
-my_formula <- as.formula(paste(name_of_target, paste(vars, collapse=" + "), sep=" ~ "))
-modelling_start = Sys.time() # start to measure time
-tree_model <- rpart(formula = my_formula, data = train,
-                    method = "class", control = rpart.control(cp = 0))
-y_pred <- as.data.frame(predict(tree_model, test, type = "class"))
-modelling_stop = Sys.time() # end measuring time
-# calculating metrics
-confusion_matrix <- get_confusion_matrix(test[[name_of_target]], y_pred[,1])
-accuracy_v <- accuracy(confusion_matrix)
-precision_v <- precision(confusion_matrix)
-recall_v <- recall(confusion_matrix)
-f1_v <- f1(confusion_matrix)
-classification_report <- data.frame(accuracy_v, precision_v,
-                                    recall_v, f1_v)
-colnames(classification_report) <- c("accuracy", "precision",
-                                     "recall", "f1")
-dataset_list$dataset <- NULL
-# in future maybe return all dataset_list ?
-# for now stick with readability
-imp_method_name <- deparse(substitute(imputation_fun))
-return(list( dataset_id = dataset_list$id, 
-             imp_method = imp_method_name,
-             confusion_matrix = confusion_matrix,
-             classification_report = classification_report,
-             imputation_time = imputation_stop - imputation_start,
-             modelling_time = modelling_stop - modelling_start))
-}
-data_all <- read_all_datasets()
-# imputations and targets preparation
-imputations <- list(imputation_fun_vim, imputation_fun_missForest,
-                  imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
-targets <- lapply(data_all, function(d){d$target})
-```
-
-Here are the results:
-
-```{r warning=FALSE, cache = TRUE}
-  library(knitr)
-  results <- rep(list(0), 5)
-  results[[1]] <- readRDS('./part_results/res1_new.rds')
-  results[[2]] <- readRDS('./part_results/res2_new.rds')
-  results[[3]] <- readRDS('./part_results/res3_new.rds')
-  results[[4]] <- readRDS('./part_results/res4_new.rds')
-  results[[5]] <- readRDS('./part_results/res5_new.rds')
-
-  result_table <- function(ds_it){
-    d <- data.frame(matrix(ncol = 7, nrow = 0))
-    colnames(d) <- c("imputation", "accuracy", "precision", "recall", "f1", "imp_time", "mod_time")
-    for (i in 1:length(imputations)){
-      if (length(results[[i]][[ds_it]]) != 1){ # length equal 1 means "ERROR" was generated in partial results 
-        d[i,] <- c(results[[i]][[ds_it]]$imp_method,
-               round(as.numeric(results[[i]][[ds_it]]$classification_report$accuracy),3),   
-               round(as.numeric(results[[i]][[ds_it]]$classification_report$precision),3),
-               round(as.numeric(results[[i]][[ds_it]]$classification_report$recall),3),
-               round(as.numeric(results[[i]][[ds_it]]$classification_report$f1),3),
-               round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
-               # @TODO konwersja jednostek, as.numeric nie rozroznia jednostek czasu
-               round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
-      }
-      else{ # for error input imputation name
-        if (i==3){
-          d[i,1] <- "imputation_fun_mice" 
-        }
-        if (i==4){
-          d[i,1] <- "imputation_fun_missForest" 
-        }
-        if (i==5){
-          d[i,1] <- "imputation_fun_vim" 
-        }
-      }
-    }
-    d
-  }
-
-  # change i bound to number of datasets 
-  #for (i in 1:length(results[[1]])){ # won't use it, because the output looks bad
-  #  kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
-  #}
-
-  kable(result_table(1), caption = paste("Dataset id: ", results[[2]][[1]]$dataset_id, sep = ""))
-  kable(result_table(2), caption = paste("Dataset id: ", results[[2]][[2]]$dataset_id, sep = ""))
-  kable(result_table(3), caption = paste("Dataset id: ", results[[2]][[3]]$dataset_id, sep = ""))
-  kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
-  kable(result_table(5), caption = paste("Dataset id: ", results[[2]][[5]]$dataset_id, sep = ""))
-  kable(result_table(6), caption = paste("Dataset id: ", results[[2]][[6]]$dataset_id, sep = ""))
-  kable(result_table(7), caption = paste("Dataset id: ", results[[2]][[7]]$dataset_id, sep = ""))
-  kable(result_table(8), caption = paste("Dataset id: ", results[[2]][[8]]$dataset_id, sep = ""))
-  kable(result_table(9), caption = paste("Dataset id: ", results[[2]][[9]]$dataset_id, sep = ""))
-  kable(result_table(10), caption = paste("Dataset id: ", results[[2]][[10]]$dataset_id, sep = ""))
-  kable(result_table(11), caption = paste("Dataset id: ", results[[2]][[11]]$dataset_id, sep = ""))
-  kable(result_table(12), caption = paste("Dataset id: ", results[[2]][[12]]$dataset_id, sep = ""))
-  kable(result_table(13), caption = paste("Dataset id: ", results[[2]][[13]]$dataset_id, sep = ""))
-  kable(result_table(14), caption = paste("Dataset id: ", results[[2]][[14]]$dataset_id, sep = ""))
-```
+Also, it is worth mentioning that removing rows computes a different dataset compared to other techniques.
+Because of that one needs to be cautious while analysing and comparing remove_rows to other imputation methods.
+For one's own research it is probably worth replacing removing rows with removing columns for datasets with few features containing missing values.
+Here are the results. NA indicates that method returned error:
 
@@
 
-```{r imputation_times_counting, cache = TRUE}
-ndatasets <- length(results[[1]])
-nimputations <- length(imputations)
-imputation_names <- c("remove", "median", "mice", "missForest", "VIM")
-dataset_ids <- rep(NULL, ndatasets)
-for(i in 1:ndatasets){
-  dataset_ids[i] <- results[[2]][[i]]$dataset_id
-}
-
-imputation_times <- data.frame(rep(0, ndatasets))
-for (imputation_id in 1:nimputations){
-  imp_id_times <- rep(NULL, ndatasets)
-
-  for(dataset_id in 1:ndatasets){
-    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
-      imp_time <- results[[imputation_id]][[dataset_id]]$imputation_time
-      imp_time <- as.numeric(imp_time, units="secs")
-      imp_id_times[dataset_id] <- imp_time
-    }
-  }
-  imputation_times <- cbind(imputation_times, imp_id_times)
-}
-
-imputation_times[1] <- dataset_ids
-colnames(imputation_times) <- c("dataset_id", imputation_names)
-imputation_times$dataset_id <- as.factor(imputation_times$dataset_id)
-knitr::kable(imputation_times)
-```
-
 Looks good! Now it's visualization time.
-
-```{r imputation_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
-library(ggplot2)
-# install.packages("reshape") # if not installed
-library(reshape)
-
-# in order to connect colors and shapes in legend see:
-# https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend
-
-colors <- c("remove" = "orange", "median" = "cyan", "mice" = "green",
-            "missForest" = "red", "VIM" = "blue")
-imptimes <- melt(imputation_times)
-
-require(scales)
-ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' imputations times") +
-  theme_gray() +
-  scale_y_continuous(trans = log2_trans(),
-    breaks = trans_breaks("log2", function(x) 2^x),
-    labels = trans_format("log2", math_format(2^.x))) +
-  labs(x = "Dataset Id",
-       y = "Imputation time [in seconds]",
-       color = "Legend") +
-  scale_color_manual(name = "imputation name", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-```
 
@@
 What do the numbers say?
-
-```{r imputation_times_measures, cache = TRUE}
-means <- rep(NA, nimputations)
-medians <- rep(NA, nimputations)
-stDeviations <- rep(NA, nimputations)
-
-for(i in (1:nimputations)){
-  means[i] <- mean(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
-  means[i] <- round(means[i], 3)
-  medians[i] <- median(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
-  medians[i] <- round(medians[i], 3)
-  stDeviations[i] <- sd(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
-  stDeviations[i] <- round(stDeviations[i], 3)
-}
-
-imputation_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
-knitr::kable(imputation_times_measures)
-```
 
@@
 
-```{r imputation_times_measures_little, cache = TRUE}
-dataset_ids_fullimp <- c(2, 5, 8, 11, 13, 14)
-ndatasets_fullimp <- length(dataset_ids_fullimp)
-
-means <- rep(NA, nimputations)
-medians <- rep(NA, nimputations)
-stDeviations <- rep(NA, nimputations)
-
-for(i in 1:nimputations){
-  means[i] <- mean(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
-  means[i] <- round(means[i], 3)
-  medians[i] <- median(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
-  medians[i] <- round(medians[i], 3)
-  stDeviations[i] <- sd(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
-  stDeviations[i] <- round(stDeviations[i], 3)
-}
-
-imputation_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
-knitr::kable(imputation_times_measures_fullimp)
-```
-
 Now we have better results to compare.
@@
 As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.
-
-## Comparing modeling times
-
-Let's also take a look at modeling times.
-
-```{r modeling_times_counting, cache = TRUE}
-modeling_times <- data.frame(rep(0, ndatasets))
-
-for (imputation_id in 1:nimputations){
-  mod_id_times <- rep(NULL, ndatasets)
-
-  for(dataset_id in 1:ndatasets){
-    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
-      mod_time <- results[[imputation_id]][[dataset_id]]$modelling_time
-      mod_time <- as.numeric(mod_time, units="secs")
-      mod_id_times[dataset_id] <- mod_time
-    }
-  }
-  modeling_times <- cbind(modeling_times, mod_id_times)
-}
-
-modeling_times[1] <- dataset_ids
-colnames(modeling_times) <- c("dataset_id", imputation_names)
-imputation_times$dataset_id <- as.factor(modeling_times$dataset_id)
-knitr::kable(modeling_times)
-```
-
-Very interesting!
-The longest modeling time took one of the largest sets...
-on deficiencies supplemented by median.
-And it took apparently longer than other modeling!
-could a random tree have more to do with a lot of the same values?
-Probably yes.
-Let's visualize our results.
-
-```{r modeling_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
-modeling_times$dataset_id <- as.factor(modeling_times$dataset_id)
-modtimes <- melt(modeling_times)
-
-require(scales)
-ggplot(data=modtimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' modeling times") +
-  theme_light() +
-  scale_y_continuous(trans = log2_trans(),
-    breaks = trans_breaks("log2", function(x) 2^x),
-    labels = trans_format("log2", math_format(2^.x))) +
-  labs(x = "Dataset Id",
-       y = "Modeling time [in seconds]",
-       color = "Legend") +
-  scale_color_manual(name = "imputation name", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-```
-
-Here the results seem to be way more scattered. Let's check on measures also.
-
-```{r modeling_times_measures, cache = TRUE}
-means <- rep(NA, nimputations)
-medians <- rep(NA, nimputations)
-stDeviations <- rep(NA, nimputations)
-
-for(i in (1:nimputations)){
-  means[i] <- mean(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
-  means[i] <- round(means[i], 3)
-  medians[i] <- median(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
-  medians[i] <- round(medians[i], 3)
-  stDeviations[i] <- sd(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
-  stDeviations[i] <- round(stDeviations[i], 3)
-}
-
-modeling_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
-knitr::kable(modeling_times_measures)
-```
-
-The powerful but also fast tree classifier from *rpart* (both looking at the median and average) counted everything in less than a second.
-The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds.
-You can also see very similar all three average measures for these two, probably due to the operation of *VIM* as a result giving similar effects as the *median*.
-Interestingly, unbeatably the fastest tree was built for the *mice* algorithm ...
-but wait!
-Maybe we'd better compare only the imputations for those sets for which we've done everything.
-
-```{r modeling_times_measures_little, cache = TRUE}
-means <- rep(NA, nimputations)
-medians <- rep(NA, nimputations)
-stDeviations <- rep(NA, nimputations)
-
-for(i in 1:nimputations){
-  means[i] <- mean(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
-  means[i] <- round(means[i], 3)
-  medians[i] <- median(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
-  medians[i] <- round(medians[i], 3)
-  stDeviations[i] <- sd(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
-  stDeviations[i] <- round(stDeviations[i], 3)
-}
-
-modeling_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
-knitr::kable(modeling_times_measures_fullimp)
-```
-
-On the attached non-manipulated table, all average measures are almost identical.
-The conclusion is that the method of imputation probably has almost no effect on the modeling time - although it is difficult to say, given that we had a fairly small number of sets to compare and we have values in the order of hundredths of a second - probably just the operation of the computer pretty badly distorts the result.
 
@@
 
-```{r measures_tables, cache = TRUE}
-bin_ids <- c(1, 3, 4, 5, 6, 8, 9, 10, 11, 12, 13, 14)
-
-accuracies <- data.frame(rep(0, ndatasets))
-precisions <- data.frame(rep(0, ndatasets))
-recalls <- data.frame(rep(0, ndatasets))
-f1s <- data.frame(rep(0, ndatasets))
-
-for (imputation_id in 1:nimputations){
-  acc_id <- rep(NA, ndatasets)
-  pre_id <- rep(NA, ndatasets)
-  rec_id <- rep(NA, ndatasets)
-  f1_id <- rep(NA, ndatasets)
-
-  for(dataset_id in bin_ids){
-    if("classification_report" %in% names(results[[imputation_id]][[dataset_id]])){
-      accuracy <- results[[imputation_id]][[dataset_id]]$classification_report$accuracy
-      acc_id[dataset_id] <- accuracy
-      precision <- results[[imputation_id]][[dataset_id]]$classification_report$precision
-      pre_id[dataset_id] <- precision
-      recall <- results[[imputation_id]][[dataset_id]]$classification_report$recall
-      rec_id[dataset_id] <- recall
-      f1 <- results[[imputation_id]][[dataset_id]]$classification_report$f1
-      f1_id[dataset_id] <- f1
-    }
-  }
-  accuracies <- cbind(accuracies, acc_id)
-  precisions <- cbind(precisions, pre_id)
-  recalls <- cbind(recalls, rec_id)
-  f1s <- cbind(f1s, f1_id)
-}
-
-accuracies[1] <- dataset_ids
-colnames(accuracies) <- c("dataset_id", imputation_names)
-accuracies$dataset_id <- as.factor(accuracies$dataset_id)
-accuracies <- accuracies[-c(2, 7), ] # delete  non-binary classification datasets
-knitr::kable(accuracies, caption = "Accuracies")
-
-precisions[1] <- dataset_ids
-colnames(precisions) <- c("dataset_id", imputation_names)
-precisions$dataset_id <- as.factor(precisions$dataset_id)
-precisions <- precisions[-c(2, 7), ] # delete  non-binary classification datasets
-knitr::kable(precisions, caption = "Precisions")
-
-recalls[1] <- dataset_ids
-colnames(recalls) <- c("dataset_id", imputation_names)
-recalls$dataset_id <- as.factor(recalls$dataset_id)
-recalls <- recalls[-c(2, 7), ] # delete  non-binary classification datasets
-knitr::kable(recalls, caption = "Recalls")
-
-f1s[1] <- dataset_ids
-colnames(f1s) <- c("dataset_id", imputation_names)
-f1s$dataset_id <- as.factor(f1s$dataset_id)
-f1s <- f1s[-c(2, 7), ] # delete  non-binary classification datasets
-knitr::kable(f1s, caption = "F1s")
-```
-
 Much data!!
@@
 Naturally - we'll visualize it.
-
-```{r measures_visualizations, message = FALSE, warning = FALSE, cache = TRUE}
-accMelt <- melt(accuracies)
-accPlot <- ggplot(data=accMelt, aes(x = dataset_id, y = value, color = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' accuracies") +
-  ylim(0, 1) +
-  labs(x = "Dataset Id",
-       y = "Accuracy",
-       color = "Legend") +
-  scale_color_manual(name = "accuracy", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-accPlot
-
-preMelt <- melt(precisions)
-prePlot <- ggplot(data=preMelt, aes(x = dataset_id, y = value, color = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' precisions") +
-  ylim(0, 1) +
-  labs(x = "Dataset Id",
-       y = "Precision",
-       color = "Legend") +
-  scale_color_manual(name = "precision", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-prePlot
-
-recMelt <- melt(recalls)
-recPlot <- ggplot(data=recMelt, aes(x = dataset_id, y = value, color = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' recalls") +
-  ylim(0, 1) +
-  labs(x = "Dataset Id",
-       y = "Recall",
-       color = "Legend") +
-  scale_color_manual(name = "recall", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-recPlot
-
-f1Melt <- melt(f1s)
-f1Plot <- ggplot(data=f1Melt, aes(x = dataset_id, y = value, color = variable)) +
-  geom_point() + 
-  ggtitle("Datasets' f1s") +
-  ylim(0, 1) +
-  labs(x = "Dataset Id",
-       y = "F1",
-       color = "Legend") +
-  scale_color_manual(name = "f1", values = colors) +
-  theme(legend.position="right",
-        plot.title = element_text(hjust = 0.5))
-f1Plot
-
-library(gridExtra)
-grid.arrange(accPlot +
-               ggtitle("Accuracies") +
-               theme(legend.position = "none",
-                axis.line.x = element_blank(),
-                axis.ticks.x = element_blank(),
-                axis.text.x = element_blank(),
-                axis.title = element_blank()),
-              prePlot +
-               ggtitle("Precisions") +
-               theme(legend.position = "none",
-                axis.line.x = element_blank(),
-                axis.ticks.x = element_blank(),
-                axis.text.x = element_blank(),
-                axis.title = element_blank()),
-              recPlot +
-               ggtitle("Recalls") +
-               theme(legend.position = "none",
-                axis.line.x = element_blank(),
-                axis.ticks.x = element_blank(),
-                axis.text.x = element_blank(),
-                axis.title = element_blank()),
-             f1Plot +
-               ggtitle("F1s") +
-               theme(legend.position = "none",
-                axis.line.x = element_blank(),
-                axis.ticks.x = element_blank(),
-                axis.text.x = element_blank(),
-                axis.title = element_blank()),
-             nrow = 2,
-             top = "Measures values for different datasets' imputations")
-```
 
@@
 
-```{r measures_measures, cache = TRUE}
-accMeans <- rep(NA, nimputations)
-accMedians <- rep(NA, nimputations)
-accStDeviations <- rep(NA, nimputations)
-
-preMeans <- rep(NA, nimputations)
-preMedians <- rep(NA, nimputations)
-preStDeviations <- rep(NA, nimputations)
-
-recMeans <- rep(NA, nimputations)
-recMedians <- rep(NA, nimputations)
-recStDeviations <- rep(NA, nimputations)
-
-f1sMeans <- rep(NA, nimputations)
-f1sMedians <- rep(NA, nimputations)
-f1sStDeviations <- rep(NA, nimputations)
-
-for(i in (1:nimputations)){
-    accMeans[i] <- mean(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
-    accMeans[i] <- round(accMeans[i], 3)
-    accMedians[i] <- median(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
-    accMedians[i] <- round(accMedians[i], 3)
-    accStDeviations[i] <- sd(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
-    accStDeviations[i] <- round(stDeviations[i], 3)
-
-    preMeans[i] <- mean(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
-    preMeans[i] <- round(preMeans[i], 3)
-    preMedians[i] <- median(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
-    preMedians[i] <- round(preMedians[i], 3)
-    preStDeviations[i] <- sd(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
-    preStDeviations[i] <- round(preStDeviations[i], 3)
-
-    recMeans[i] <- mean(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
-    recMeans[i] <- round(recMeans[i], 3)
-    recMedians[i] <- median(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
-    recMedians[i] <- round(recMedians[i], 3)
-    recStDeviations[i] <- sd(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
-    recStDeviations[i] <- round(recStDeviations[i], 3)
-
-    f1sMeans[i] <- mean(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
-    f1sMeans[i] <- round(f1sMeans[i], 3)
-    f1sMedians[i] <- median(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
-    f1sMedians[i] <- round(f1sMedians[i], 3)
-    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
-    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
-}
-
-measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
-                                preMeans, preMedians, preStDeviations,
-                                recMeans, recMedians, recStDeviations,
-                                f1sMeans, f1sMedians, f1sStDeviations)
-
-knitr::kable(measures_measures, content = "Model measures' concentrations' measures")
-```
-
 Looking at the table quickly, we can probably set our podium!!
@@
 This is necessary, because if not for the sets in which, naturally, regardless of the model or imputation we get very high or very low results generate a lot of chaos.
-
-```{r correct_measures_measures, cache = TRUE}
-ids_all_model_measures <- c(4, 6, 9, 11, 12)
-
-accMeans <- rep(NA, nimputations)
-accMedians <- rep(NA, nimputations)
-accStDeviations <- rep(NA, nimputations)
-
-preMeans <- rep(NA, nimputations)
-preMedians <- rep(NA, nimputations)
-preStDeviations <- rep(NA, nimputations)
-
-recMeans <- rep(NA, nimputations)
-recMedians <- rep(NA, nimputations)
-recStDeviations <- rep(NA, nimputations)
-
-f1sMeans <- rep(NA, nimputations)
-f1sMedians <- rep(NA, nimputations)
-f1sStDeviations <- rep(NA, nimputations)
-
-for(i in 1:nimputations){
-    accMeans[i] <- mean(accuracies[imputation_names[i]][ids_all_model_measures,])
-    accMeans[i] <- round(accMeans[i], 3)
-    accMedians[i] <- median(accuracies[imputation_names[i]][ids_all_model_measures,])
-    accMedians[i] <- round(accMedians[i], 3)
-    accStDeviations[i] <- sd(accuracies[imputation_names[i]][ids_all_model_measures,])
-    accStDeviations[i] <- round(stDeviations[i], 3)
-
-    preMeans[i] <- mean(precisions[imputation_names[i]][ids_all_model_measures,])
-    preMeans[i] <- round(preMeans[i], 3)
-    preMedians[i] <- median(precisions[imputation_names[i]][ids_all_model_measures,])
-    preMedians[i] <- round(preMedians[i], 3)
-    preStDeviations[i] <- sd(precisions[imputation_names[i]][ids_all_model_measures,])
-    preStDeviations[i] <- round(preStDeviations[i], 3)
-
-    recMeans[i] <- mean(recalls[imputation_names[i]][ids_all_model_measures,])
-    recMeans[i] <- round(recMeans[i], 3)
-    recMedians[i] <- median(recalls[imputation_names[i]][ids_all_model_measures,])
-    recMedians[i] <- round(recMedians[i], 3)
-    recStDeviations[i] <- sd(recalls[imputation_names[i]][ids_all_model_measures,])
-    recStDeviations[i] <- round(recStDeviations[i], 3)
-
-    f1sMeans[i] <- mean(f1s[imputation_names[i]][ids_all_model_measures,])
-    f1sMeans[i] <- round(f1sMeans[i], 3)
-    f1sMedians[i] <- median(f1s[imputation_names[i]][ids_all_model_measures,])
-    f1sMedians[i] <- round(f1sMedians[i], 3)
-    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][ids_all_model_measures,])
-    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
-}
-
-correct_measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
-                                preMeans, preMedians, preStDeviations,
-                                recMeans, recMedians, recStDeviations,
-                                f1sMeans, f1sMedians, f1sStDeviations)
-
-knitr::kable(correct_measures_measures, content = "CORRECT model measures' concentrations' measures")
-```
 
@@
 Let's visualize dataset's model results measures.
-
-```{r model_measures_plots, message = FALSE, cache = TRUE}
-colors2 <- c("mean" = "blue", "median" = "orange")
-
-acc_measures <- correct_measures_measures %>% select(imputation_names, accMeans, accMedians)
-colnames(acc_measures) <- c("imputation_name", "mean", "median")
-accMelt <- melt(acc_measures)
-accuracyPlot <- ggplot(data = accMelt, aes(x = imputation_name, y = value, color = variable)) +
-  geom_point(size = 3) + 
-  ylim(0.75, 1) +
-  theme_light() +
-  ggtitle("Accuracies") +
-  labs(color = "Legend") +
-  scale_color_manual(name = "measure", values = colors2) +
-  theme(legend.position="None",
-        plot.title = element_text(hjust = 0.5),
-        axis.title = element_blank())
-
-pre_measures <- correct_measures_measures %>% select(imputation_names, preMeans, preMedians)
-colnames(pre_measures) <- c("imputation_name", "mean", "median")
-preMelt <- melt(pre_measures)
-precisionPlot <- ggplot(data = preMelt, aes(x = imputation_name, y = value, color = variable)) +
-  geom_point(size = 3) + 
-  ylim(0.75, 1) +
-  theme_light() +
-  ggtitle("Precisions") +
-  labs(color = "Legend") +
-  scale_color_manual(name = "measure", values = colors2) +
-  theme(legend.position="None",
-        plot.title = element_text(hjust = 0.5),
-        axis.title = element_blank())
-
-rec_measures <- correct_measures_measures %>% select(imputation_names, preMeans, preMedians)
-colnames(rec_measures) <- c("imputation_name", "mean", "median")
-recMelt <- melt(rec_measures)
-recallPlot <- ggplot(data = recMelt, aes(x = imputation_name, y = value, color = variable)) +
-  geom_point(size = 3) + 
-  ylim(0.75, 1) +
-  theme_light() +
-  ggtitle("Recalls") +
-  labs(x = "Imputation name",
-       y = "Recall",
-       color = "Legend") +
-  scale_color_manual(name = "measure", values = colors2) +
-  theme(legend.position="None",
-        plot.title = element_text(hjust = 0.5),
-        axis.title = element_blank())
-
-f1_measures <- correct_measures_measures %>% select(imputation_names, f1sMeans, f1sMedians)
-colnames(f1_measures) <- c("imputation_name", "mean", "median")
-f1mMelt <- melt(f1_measures)
-f1sPlot <- ggplot(data = f1mMelt, aes(x = imputation_name, y = value, color = variable)) +
-  geom_point(size = 3) + 
-  ylim(0.75, 1) +
-  theme_light() +
-  ggtitle("F1s") +
-  labs(x = "Imputation name",
-       y = "f1",
-       color = "Legend") +
-  scale_color_manual(name = "measure", values = colors2) +
-  theme(legend.position="None",
-        plot.title = element_text(hjust = 0.5),
-        axis.title = element_blank())
-
-grid.arrange(accuracyPlot, precisionPlot,
-             recallPlot, f1sPlot, nrow = 2,
-             top = "Means (blue) and medians (orange) of datasets' imputations' modeling results")
-```
 
@@
 
-```{r imputation_times_measures_medians, cache = TRUE}
-medians_measures_times <- correct_measures_measures %>% select(imputation_names, accMedians, preMedians, recMedians, f1sMedians)
-medians_measures_times <- cbind(medians_measures_times, imputation_times_measures_fullimp$medians)
-colnames(medians_measures_times) <- c("imputation name", "accuracy", "precision", "recall", "f1", "imputation mean time")
-knitr::kable(medians_measures_times, content = "Mean imputation time and mean model result's measures of imp methods")
-```
-
 Looking at the results, *VIM* and *missForest* give quite similar numbers in terms of satisfaction, but VIM is much faster - for data sets for which each imputation has passed, it managed in six times shorter period of time than *missForest* (and also of course much worse *mice*).
@@
 
-## Conclusion
+# Conclusion
 
@@
 Nevertheless, above code is certainly very valuable and for more data sets it could confirm one or another belief - therefore we can probably say that we have achieved scientific success!
+
+## References
+
+*^1^mice: Multivariate Imputation by Chained Equations, package created by Stef van Buuren, Karin Groothuis-Oudshoorn, Gerko Vink, Rianne Schouten, Alexander Robitzsch, Lisa Doove, Shahab Jolani, Margarita Moreno-Betancur, Ian White, Philipp Gaffert, Florian Meinfelder, Bernie Gray*
+
+*^2^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
+
+*^3^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
+
+*^4^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*

### 2020-05-20 rankingi.R
rankingi

@@ -1,0 +1,54 @@
+library(dplyr)
+final_results2 <- readRDS('./final_results_2.RDS')
+f2 <- unlist(final_results2, recursive = FALSE)
+
+times <- lapply(f2, function(x){list( x$dataset_id,
+                                      #as.numeric(x$imputation_time, units = 'secs'), 
+                                      as.character(x$imp_method), 
+                                      as.character(x$mod_mehtod), 
+                                      as.numeric(x$classification_report$f1), 
+                                      as.numeric(x$classification_report$mcc))})
+
+df <- do.call(rbind, times) %>% as.data.frame()
+
+#colnames(df) <- c("time", "imp", "model", "f1", "mcc")
+colnames(df) <- c("id","imp", "model", "f1", "mcc")
+
+#df$time <- as.numeric(df$time)
+df$id <- as.integer(df$id)
+df$imp <- as.character(df$imp)
+df$model <- as.character(df$model)
+df$f1 <- as.numeric(df$f1)
+df$mcc <- as.numeric(df$mcc)
+
+rankings <- df %>% group_by(id, model) %>% mutate(r_f1 = rank(f1), r_mcc = rank(mcc))
+
+melted <- reshape::melt(rankings[,c("imp", "r_f1", "r_mcc")] %>% as.data.frame)
+
+library(ggplot2)
+
+# wykres nie uzględnia NA w zbiorach, które nie zostały policzone
+# nie ma dużych różnic między metryką mcc a f1
+ggplot(melted, aes(x = imp, y = value, color = variable)) +
+  geom_boxplot() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# uśrednione wyniki
+rankings %>% group_by(imp) %>% transmute(m_r_f1 = mean(r_f1),m_r_mcc = mean(r_mcc)) %>% unique()
+
+# dorzucmy NA tam, gdzie nie przeszły imputacje
+# 14 zbiorow, 6 imp, 5 modeli. Każdy imp powinien mieć 14*5 obserwacji - jeśli nie ma - wypełniamy braki NA
+
+library(tidyr)
+cmp <- df %>% 
+  complete(id, imp, model)
+
+rankings <- cmp %>% group_by(id, model) %>% mutate(r_f1 = rank(f1), r_mcc = rank(mcc))
[…]

### 2020-05-22 article.Rmd
Add files via upload

@@
 ---
-title: "Jaki tytuł??"
+title: "The Hajada Imputation Test"
 author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
@@
 
-Choosing the best-suited imputation is the daily dilemma of every data scientist.
-Some of them believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.
+There are many different ways of dealing with missing values.
+Depending on their quantity and properties, various methods turn out to be the best.
+The fact is, there is no one universal best method of imputation, and different problems require various solutions.
+Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us, or when with due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
 
-In this document we will try to find the one, objectively finest imputation method by testing six of popular ones on forteen specially selected data sets prepared for these operations.
+In this chapter which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
+Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
+For this purpose, we'll test the modified collections for getting rid of empty values on several machine learning classification algorithms, and then considering the selected quality measures after comparing the results, we'll get the desired view.
+Due to the need to obtain a clear and interpretable result, we have limited ourselves to the problem of binary classification.
 
@@
 
+In statistics, imputation is the process of replacing missing data with substituted values.
+Over the years, humanity has met many different methods of it, including the simple and instinctive ones but also more advanced and hard to be easily explained.
+Choosing the best-suited imputation method for the dataset with missing values is still the daily dilemma of every data scientist and there isn't one universally recognized best model.
+Some scientists believe the crux lies in the most advanced and sophisticated algorithms, the others trust the simplest of all possible.
+
 ## Imputation functions
 
+Six different methods have been taken into consideration.
+Their selection is well thought out - they are all widely known in the world of Data Science, and at the same time they differ significantly in approach, implementation, concept and results.
+Comparing them not only results in a good view of the strictly indicated implementations, but also what result subsequent approaches result in.
+
 We'll look at the three simple imputation methods and three more advanced algorithms offered by popular packages.
+The clash of implementations with such a degree of diversity of complexity will give a clear message whether it pays to reach because the advanced algorithm from the package prepared by professionals for many hours conducted on many tests or can, however, stay with minimalism and write easy code solving the problem in a trivial way.
 
-TODO: czemu one; lanie wody o tym, że są fajne bo różne i badamy różne spojrzenia
+Methods compared in subsequent stages are:
 
-1. *mode / median replace* - basic method which puts in a place of missing data median cells from their columns for numerical values and dominants for categorical ones.
+1. *mode / median replace* - basic process which puts in a place of missing data median cells from their columns for numerical values and dominants for categorical ones.
 2. *remove rows* - trivial solution of removing rows containing any missing data.
@@
 4.
-*mice* ^1^ - method creating multiple imputations for multivariate missing data, based on Fully Conditional Specification, where each incomplete variable is imputed by a separate model.
-Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
-5. *VIM* ^2^ - standard k-nearest neighbors algorithm taken from the library
-6. *missForest* ^3^ - simple imputation based on random forest offered by the package
+*mice* ^1^ - advanced method creating multiple imputations for multivariate missing data, based on *Fully Conditional Specification* ^2^, where each incomplete variable is imputed by a separate model.
+Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets.
+5. *VIM* ^3^ - standard k-nearest neighbors algorithm taken from the library.
+6. *missForest* ^4^ - simple imputation based on random forest offered by the package.
 
@@
 
-The fourteen data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research.
-They can be found under the following identifiers with the following names:
+The fourteen data frames were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:
 
@@
 
-TODO: tylko na sześciu wszystko się udało
+The six imputations mentioned earlier were released on all fourteen of them, but only on six imputations was successful in their entirety.
 
@@
 
-The results of our work will be checked using the implemented classification tree algorithm from the *rpart ^4^* package.
+The results of our work will be checked using the implemented classification tree algorithm from the *rpart ^5^* package.
 We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a decent model.
@@
 
-*^2^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
+*^2^Fully conditional specification: Imputes multivariate missing data on a variable-by-variable basis (Van Buuren et al. 2006; Van Buuren 2007a).*
 
-*^3^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
+*^3^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
 
-*^4^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*
+*^4^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
+
+*^5^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*

### 2020-05-22 article.Rmd
Wieczorne robótki

@@
 
-TODO: ładniejszy format
-
-The fourteen data frames were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:
-
-* 1590 - *adult*
-* 188 - *eucalyptus*
-* 23381 - *dresses-sales*
-* 29 - *credit-approval*
-* 38 - *sick*
-* 40536 - *SpeedDating*
-* 41278 - *okcupid-stem*
-* 56 - *vote*
-* 6332 - *cylinder-bands*
-* 1018 - *ipums_la_99-small*
-* 27 - *colic*
-* 4 - *labor*
-* 55 - *hepatitis*
-* 944 - *echoMonths*
+The fourteen data frames were taken from **OpenML100 collection** and were corrected specifically for this research.
+Both small and simple frames were chosen, as well as more complex and containing a lot of information, difficult to explain or present.
+They can be found under the following identifiers with the following names: 1590 - *adult*, 188 - *eucalyptus*, 23381 - *dresses-sales*, 29 - *credit-approval*, 38 - *sick*, 40536 - *SpeedDating*, 41278 - *okcupid-stem*, 56 - *vote*, 6332 - *cylinder-bands*, 1018 - *ipums_la_99-small*, 27 - *colic*, 4 - *labor*, 55 - *hepatitis* and 944 - *echoMonths*.
 
@@
 
-TODO: pięć metod, ich krótki opis i również krótkie stwierdzenie, że sa fajne bo różne i wgl.
+In order to compare the quality of the imputed data and see how this supplementation of deficiencies works in practice, five binary classification algorithms have been selected.
+Her choice was made after careful analysis and extensive discussion, in order to find models that are both widely known and used, and also apply to different approaches and give reasonably distinguishable results using different techniques.
+The final choice fell on:
+
+* *classification tree from rpart^5^* - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
+* *k-Nearest Neighbors from class^6^ package* - standard knn model attributing a given observationto a target corresponding to its closest observations in space,
+* *naive Bayes from e1071^7^* - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
+* *random forest from ranger^8^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree to try to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
+* *Support Vector Machine from e1071^7^* -  a discriminative classifier formally defined by a separating hyperplane.
 
@@
 
-The results of our work will be checked using the implemented classification tree algorithm from the *rpart ^5^* package.
+The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package.
 We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a decent model.
@@
 *^5^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*
+
+*^6^class: Various functions for classification, including k-nearest neighbour, Learning Vector Quantization and Self-Organizing Maps, package created by Brian Ripley, William Venables*
+
+*^7^e1071: Misc Functions of the Department of Statistics, Probability Theory Group (Formerly: E1071), TU Wien, package created by David Meyer*
+
+*^8^ranger: A Fast Implementation of Random Forests, package created by Marvin N. Wright, Stefan Wager, Philipp Probst*

### 2020-05-28 rankingi.R
poprawione rankingi

@@ -21,7 +21,7 @@ df$model <- as.character(df$model)
 df$f1 <- as.numeric(df$f1)
 df$mcc <- as.numeric(df$mcc)
 
-rankings <- df %>% group_by(id, model) %>% mutate(r_f1 = rank(f1), r_mcc = rank(mcc))
+rankings <- df %>% group_by(id, model) %>% mutate(r_f1 = rank(-f1), r_mcc = rank(-mcc))
 
 melted <- reshape::melt(rankings[,c("imp", "r_f1", "r_mcc")] %>% as.data.frame)
 
@@ -43,7 +43,12 @@ library(tidyr)
 cmp <- df %>% 
   complete(id, imp, model)
 
-rankings <- cmp %>% group_by(id, model) %>% mutate(r_f1 = rank(f1), r_mcc = rank(mcc))
+cmp <- replace_na(cmp, list(f1 = c(0), mcc = c(-1)))
+
+#cmp$f1[is.na(cmp)] <- 0
+#cmp$mcc[is.na(cmp)] <- -1
+
+rankings <- cmp %>% group_by(id, model) %>% mutate(r_f1 = rank(-f1), r_mcc = rank(-mcc))
 melted <- reshape::melt(rankings[,c("imp", "r_f1", "r_mcc")] %>% as.data.frame)
 
 ggplot(melted, aes(x = imp, y = value, color = variable)) +

### 2020-05-29 article.Rmd
Omg miejmy to już za sobą

@@
 knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
+getwd()
 ```
@@
 
-In this chapter which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
+In this article which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
 Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
@@
 In order to read all datasets we used another function, which returned a dataframe containing all the important informations about the test matrices.
-
-TODO: trochę o metodzie, o funkcjach czytających; PARALLEL, itd
+The generated frames of imputation results and times were saved using parallel processing^9^.
 
@@
 
-In order to test the same test-train splits we'll use one random seed 1357 for all datasets.
-
-For each machine learning model after every imputation we will get the confusion matrix and the values of four basic metrics:
+In order to test the same test-train splits we used random seed 1357 for all datasets.
+For each machine learning model after every imputation was created confusion matrix and the values of four basic metrics:
 
@@
 
-These four will give us good information about the quality of imputation and give us a good comparison.
-We get simple and clear information about which observations are well classified, which ones are wrong and how they should be.
+... in the final conclusion the final ranking concerned only the last one. Additionally, Matthews correlation coefficient measures were also counted.
+
+${\displaystyle {\text{MCC}}={\frac {{\mathit {TP}}\times {\mathit {TN}}-{\mathit {FP}}\times {\mathit {FN}}}{\sqrt {({\mathit {TP}}+{\mathit {FP}})({\mathit {TP}}+{\mathit {FN}})({\mathit {TN}}+{\mathit {FP}})({\mathit {TN}}+{\mathit {FN}})}}}}$
+
+These two gave good information about the quality of imputation and a good comparison.
+I resulted in simple and clear information about which observations are well classified, which ones are wrong and how they should be.
 Considering the number of sets to consider and their different properties, even when it comes to balancing, comparing these four measures, the fact that it will show the pros and cons of given imputations, will additionally be good to compare them.
-
-TODO: MCC jeszcze, czemu f1
-
-## Imputation results
-
-The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package.
-We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a decent model.
-The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.
-
-Also, it is worth mentioning that removing rows computes a different dataset compared to other techniques.
-Because of that one needs to be cautious while analysing and comparing remove_rows to other imputation methods.
-For one's own research it is probably worth replacing removing rows with removing columns for datasets with few features containing missing values.
-Here are the results. NA indicates that method returned error:
 
@@
 
-Let's build a dataframe containing information about imputation methods.
+A dataframe containing information about imputation methods was built.
 
-Looks good! Now it's visualization time.
+![](files/imputationTimesScatter.png)
 
 Judging by the logarithmic scale, no surprise removing rows and median replenishment are definitely the fastest methods, with removing rows being several times faster.
-Looking at more advanced ones, definitely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*.
+what would you expect, also "random rows" was quite fast, but slower than the two mentioned.Looking at more advanced ones, definitely *VIM* turned up to be the fastest - probably usually something like 5-10 times more than *mice* and *missForest*.
 These last two are quite slow, with the missforest appearing to be slightly faster.
 
-What do the numbers say?
+To better view a boxplot was also created.
 
-Taking into account all the imputations which have been implemented, we can clearly see that two easy ones really stand out.
+![](files/imputationTimesBoxplot.png)
+
+Taking into account all the imputations which have been implemented, we can clearly see that two first mentioned ones really stand out.
 However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.
@@
 
+TO NIŻEJ SIĘ NIE PRZYDA CHYBA, ŚCIEK Z PD2
 Let's count these four measures again, but only for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
@@
 As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.
+NOOO CHYBA MOŻNA WYRZUCIĆ
 
@@
 *^8^ranger: A Fast Implementation of Random Forests, package created by Marvin N. Wright, Stefan Wager, Philipp Probst*
+
+*^9^parallel: Support for parallel computation, including by forking (taken from package multicore), by sockets (taken from package snow) and random-number generation*
+
+*^10^Matthews correlation coefficient: a measure of the quality of binary (two-class) classifications, introduced by biochemist Brian W. Matthews in 1975*

### 2020-05-30 article.Rmd
Sobota rano

@@
 
-In this article which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
+In this chapter which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
 Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
@@
 
-These two gave good information about the quality of imputation and a good comparison.
-I resulted in simple and clear information about which observations are well classified, which ones are wrong and how they should be.
+These two give good information about the quality of imputation and a good comparison.
+It resulted in simple and clear information about which observations are well classified, which ones are wrong and how they should be.
 Considering the number of sets to consider and their different properties, even when it comes to balancing, comparing these four measures, the fact that it will show the pros and cons of given imputations, will additionally be good to compare them.
@@
 
-Taking into account all the imputations which have been implemented, we can clearly see that two first mentioned ones really stand out.
+Taking into account all the imputations which have been implemented, it is clearly seen that two first mentioned ones really stand out.
 However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.
@@
 
-TO NIŻEJ SIĘ NIE PRZYDA CHYBA, ŚCIEK Z PD2
-Let's count these four measures again, but only for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
-
-Now we have better results to compare.
-Taking into account the mean time, definitely *missForest* is the slowest, but also its standard deviation is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
-As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.
-NOOO CHYBA MOŻNA WYRZUCIĆ
+Also was compared time for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
+Taking into account the mean time, definitely *missForest* was the slowest, but also its standard deviation seemed to be incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
+*VIM* turned out bo be better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so for small data sets there is not much difference between them, and a lot of time was definitely needed to devote to these larger data frames.
 
@@
 
-Let us draw attention to our measures. How did our four measures work on them?
+In order to compare *F1* and *MCC* measures, mean results of measures on all sets for all imputations were calculated.
+Then for each machine learning algorithm a ranking was created - so that the best result was obtained in the first, second - 2nd, etc.
+The results came out as shown by this boxplot:
 
-Attention!
-We will not consider sets with id 41278 and 188 in the next steps - they are non-binary classifications, so they are not affected by accuracy, precision, recall and f1.
-However, from the remaining twelve collections - we will be able to draw many valuable conclusions.
+![](files/rankingAll.png)
 
-Much data!!
+It might seem that the results of the methods are very close to each other, but the above visualization takes into account all measures - to get the correct comparison there is a need to make a ranking only for the sets for which all imputations have passed.
+And it looks like the picture below.
 
-Naturally - we'll visualize it.
+![](files/rankingGood.png)
 
-... where colors of points represent: <p style="color:orange">*1. **remove***</p>
-<p style="color:cyan">*2. **median***</p><p style="color:green">*3. **mice***</p>
-<p style="color:red">*4. **missForest***</p><p style="color:blue">*5. **VIM***</p>
-
-Looking only at these plots, it is difficult to come up with something at first glance - for each time I come to some conclusion, results for a few other datasets that I did not pay attention to deny it.
-For sure we can already say that the scatter of results is rather large and analyzing the measures of concentration, we probably will not clearly indicate the best or the worst imputation.
-
-Visualizations often help and ever give answers to the questions, but this time the best option will be definitely to calculate some measures.
-What means, medians and standard deviations give these measures?
-
-Looking at the table quickly, we can probably set our podium!!
-*MissForest* becomes the big winner just before ex aequo in second place: *mice* and *VIM*.
-At the bottom we have *median*, slightly overtaken by *remove_rows*?
-Well no!
-Remember that we should not take datasets for such comparison, for which we have only partial results.
-They probably strongly disturb the information about the data.
-The table above is not useless, because in the end we have means, medians and standard deviations of everything we could; but in order to be able to compare them without remorse, we should take only these five - which have complete information - we will calculate again the measures for datasets with ids 38, 56, 27, 55 and 944.
-This is necessary, because if not for the sets in which, naturally, regardless of the model or imputation we get very high or very low results generate a lot of chaos.
-
-So now we can make some conclusions!
-
-Having the table above, we can't say that *missForest* is definitely the best.
-Looking at the advanced ones, *missForest* and *VIM* are doing very well, almost always better than their competitors.
-*VIM* has slightly better accuracy and definitely better precision, while *missForest* dominates in the recall case and has slightly better f1.
-***Mice* is almost always the worst**.
-Even *remove_rows* is doing better than it, and the *median* sometimes catches up with missForest and mice.
-
-Standard deviations are pretty balanced and it's hard to tell anything clever about it - especially considering the small number of collections. Rather, nothing stands out.
-
-Let's visualize dataset's model results measures.
-
-I think we will all agree now - in terms of results, at least looking at these six datasets for which each imputation was successful, we can create such a podium:
-
-1. *VIM* / *missForest* (judging by our need - for *accuracy* and *precision* - *VIM*, for *recall* - *missForest*)
-3. *median*
-4. *remove_rows*
-5. *mice*
-
-Of course, the results are not very representative, given the small number of datasets for which they were compared - but still something.
-
-## Imputation time and its results
-
-We collected and summarized both results and imputation times in our collections. What are the average imputation times and the measures' medians of our imputations?
-
-Looking at the results, *VIM* and *missForest* give quite similar numbers in terms of satisfaction, but VIM is much faster - for data sets for which each imputation has passed, it managed in six times shorter period of time than *missForest* (and also of course much worse *mice*).
-The question is how much we are interested in time - it turns out that for very large data sets our model will not suffer much if we take the median instead of these two.
-At this point, you can probably make a correction of the ranking:
-
-1. *VIM*
-2. *median* / *missForest* (depending on how important time is for us)
-...
-4. *remove_rows*
-...
-5. *mice*
+Considering the imputation times and results, it can be said that the knn algorithm from the VIM package proved to be "universally" the best.
+With the original methods it falls out the fastest, and the results also have the best.
+However, it is more difficult to choose second and subsequent places - looking at the results, the methods are very quick, but they differ from the more advanced ones.
+Looking at the advanced ones, with the exception of VIM, MissForest stands out.
+However, his problem is time.
+So after VIM, it seems second best to use mean / median for a small amount of time and missForest for a large one.
+However, there was no huge difference in results between the methods - which, however, is not visible in the charts, "removeRows" is characterized by large fluctuations, which we advise against.
 
@@
 
-The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a very good package, supplementing the missing values with medians from columns is not at all is such a stupid idea as it may seem, and *mice* can be thrown in the bin.
+The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a very good package, supplementing the missing values with medians from columns is not at all is such a stupid idea as it may seem and *missForest* is also noteworthy - if we have time for running it and the dataset isn't too huge.
 Unfortunately, it is impossible to draw a wise conclusion based on such a small number of tested data - admittedly as many as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer and assumptions that respectful statistics would praise.
-Nevertheless, above code is certainly very valuable and for more data sets it could confirm one or another belief - therefore we can probably say that we have achieved scientific success!
-
+Nevertheless, above code is certainly very valuable and for more data sets it could confirm one or another belief.
 ## References
@@
 *^3^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
-
 *^4^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*

### 2020-05-30 article.Rmd
poprawki do artykułu

@@
 The fact is, there is no one universal best method of imputation, and different problems require various solutions.
-Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us, or when with due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
+Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us, or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
 
@@
 Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
-For this purpose, we'll test the modified collections for getting rid of empty values on several machine learning classification algorithms, and then considering the selected quality measures after comparing the results, we'll get the desired view.
+For this purpose, we'll test the modified collections for getting rid of missing values on several machine learning classification algorithms, and then considering the selected quality measures after comparing the results, we'll get the desired view.
 Due to the need to obtain a clear and interpretable result, we have limited ourselves to the problem of binary classification.
@@
 In statistics, imputation is the process of replacing missing data with substituted values.
-Over the years, humanity has met many different methods of it, including the simple and instinctive ones but also more advanced and hard to be easily explained.
-Choosing the best-suited imputation method for the dataset with missing values is still the daily dilemma of every data scientist and there isn't one universally recognized best model.
-Some scientists believe the crux lies in the most advanced and sophisticated algorithms, the others trust the simplest of all possible.
+Over the years, humanity has created many different methods accomplishing that, including the simple and instinctive ones but also more advanced and hard to be easily explained.
+Choosing the best-suited imputation method for the dataset with missing values is still the daily dilemma of every data scientist and there isn't one universally recognized best technique.
+Some scientists believe the crux lies in the most advanced and sophisticated algorithms, others trust the simplest of all possible.
 
@@
 Their selection is well thought out - they are all widely known in the world of Data Science, and at the same time they differ significantly in approach, implementation, concept and results.
-Comparing them not only results in a good view of the strictly indicated implementations, but also what result subsequent approaches result in.
+Comparing them not only results in a fair view of the strictly indicated implementations, but also what result subsequent approaches result in.
 
 We'll look at the three simple imputation methods and three more advanced algorithms offered by popular packages.
-The clash of implementations with such a degree of diversity of complexity will give a clear message whether it pays to reach because the advanced algorithm from the package prepared by professionals for many hours conducted on many tests or can, however, stay with minimalism and write easy code solving the problem in a trivial way.
+The clash of implementations with such a degree of diversity of complexity will give a clear message whether it pays off to reach for an advanced algorithm from the package prepared by professionals for many hours, conducted on many tests, or rather stay with minimalism and write easy code solving the problem in a trivial way.
 
@@
 2. *remove rows* - trivial solution of removing rows containing any missing data.
-3. *bogo replace* - simple algorithm replacing NA values with random numbers or factors from their columns.
+3. *'bogo' random replace* - simple algorithm replacing NA values with random numbers or factors from their columns.
 4.
@@
 5. *VIM* ^3^ - standard k-nearest neighbors algorithm taken from the library.
-6. *missForest* ^4^ - simple imputation based on random forest offered by the package.
+6. *missForest* ^4^ - imputation based on random forest offered by the package.
 
@@
 
-The six imputations mentioned earlier were released on all fourteen of them, but only on six imputations was successful in their entirety.
+The six imputations mentioned earlier were conducted on all fourteen of them, but only on six, imputations were successful in their entirety.
 
@@
 In order to compare the quality of the imputed data and see how this supplementation of deficiencies works in practice, five binary classification algorithms have been selected.
-Her choice was made after careful analysis and extensive discussion, in order to find models that are both widely known and used, and also apply to different approaches and give reasonably distinguishable results using different techniques.
+Their choice was made after careful analysis and extensive discussion, in order to find models that are both widely known and used, and also apply to different approaches and give reasonably distinguishable results using different techniques.
 The final choice fell on:
@@
 * *classification tree from rpart^5^* - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
-* *k-Nearest Neighbors from class^6^ package* - standard knn model attributing a given observationto a target corresponding to its closest observations in space,
+* *k-Nearest Neighbors from class^6^ package* - standard knn model attributing a given observation to a target corresponding to its closest observations in space,
 * *naive Bayes from e1071^7^* - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
-* *random forest from ranger^8^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree to try to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
+* *random forest from ranger^8^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree trying to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
 * *Support Vector Machine from e1071^7^* -  a discriminative classifier formally defined by a separating hyperplane.
@@
 
-... in the final conclusion the final ranking concerned only the last one. Additionally, Matthews correlation coefficient measures were also counted.
+For the final conclusion, ranking concerned only the last one. Additionally, Matthews correlation coefficient measures were also computed.
 
@@
 
-These two give good information about the quality of imputation and a good comparison.
+These two give great information about the quality of imputation and a good comparison.
 It resulted in simple and clear information about which observations are well classified, which ones are wrong and how they should be.
-Considering the number of sets to consider and their different properties, even when it comes to balancing, comparing these four measures, the fact that it will show the pros and cons of given imputations, will additionally be good to compare them.
+The main advantage of them is accounting for target variable's inbalance.
+Considering sets' different properties, even when it comes to balancing, comparing those measures, the fact that they show the pros and cons of given result will additionally be worthwhile to compare them.
 
@@
 
-Judging by the logarithmic scale, no surprise removing rows and median replenishment are definitely the fastest methods, with removing rows being several times faster.
-what would you expect, also "random rows" was quite fast, but slower than the two mentioned.Looking at more advanced ones, definitely *VIM* turned up to be the fastest - probably usually something like 5-10 times more than *mice* and *missForest*.
-These last two are quite slow, with the missforest appearing to be slightly faster.
+Judging by the logarithmic scale, there is no surprise that removing rows and mode/median replenishment are definitely the fastest methods, with removing rows being several times faster.
+what would you expect, also "random values" imputation was quite fast, but slower than the two mentioned.
+Looking at more advanced ones, definitely *VIM* turned up to be the fastest - times usually being approximately 5-10 times shorter than *missForest* and *mice*.
+These last two are quite slow, with the former appearing to be slightly faster.
 
-To better view a boxplot was also created.
+For a better view a boxplot was also created:
 
@@
 
-Taking into account all the imputations which have been implemented, it is clearly seen that two first mentioned ones really stand out.
-However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.
+Taking into account all the imputations which have been implemented, it is clearly seen that removing rows as well as replacing with mode or median really stand out.
+However, we cannot fully compare the other four - due to the fact that missForest failed on three datasets, and mice - up to seven, mostly for datasets greater in size.
 
-Considering such a large spread of data size, it is very interesting difference between the median and the average for VIM - the first is almost seven minutes, the second - barely one and a half seconds.
-In general, however, it is certainly much longer than methods for removing incomplete rows and filling with median.
+Considering such a large spread of data size, there is a very interesting difference between median and mean for VIM's  time - the first is almost seven minutes, the second - barely a second and a half.
+In general, however, it is certainly much longer than methods for removing incomplete rows and filling with mode/median.
 
-Also was compared time for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
-Taking into account the mean time, definitely *missForest* was the slowest, but also its standard deviation seemed to be incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets.
-*VIM* turned out bo be better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so for small data sets there is not much difference between them, and a lot of time was definitely needed to devote to these larger data frames.
+Also time was compared for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
+Taking into account the mean time, definitely *missForest* was the slowest, but also its standard deviation seemed to be incomparably large - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very sizeable datasets.
+*VIM* turned out to be better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so for small datasets there is not much difference between them, and a lot of time was definitely needed to devote to these larger data frames.
 
@@
 
-In order to compare *F1* and *MCC* measures, mean results of measures on all sets for all imputations were calculated.
-Then for each machine learning algorithm a ranking was created - so that the best result was obtained in the first, second - 2nd, etc.
-The results came out as shown by this boxplot:
+In order to compare *F1* and *MCC* measures, mean results on all sets for all imputations were calculated.
+Then, for each machine learning algorithm a ranking was created - so that the imputation with best results obtained the first rank, second best - 2nd, etc.
+The rank's distributions for every imputation came out as shown by the following boxplot:
 
@@
 
-It might seem that the results of the methods are very close to each other, but the above visualization takes into account all measures - to get the correct comparison there is a need to make a ranking only for the sets for which all imputations have passed.
-And it looks like the picture below.
+It might seem that the results of the methods are very close to each other, but the above visualization takes into account rankings only for successful imputations - to get the correct comparison there is a need to make a ranking with failing methods receiving the worst ranks instead of being omitted.
+The outcome was presented below:
 
@@
 Considering the imputation times and results, it can be said that the knn algorithm from the VIM package proved to be "universally" the best.
-With the original methods it falls out the fastest, and the results also have the best.
-However, it is more difficult to choose second and subsequent places - looking at the results, the methods are very quick, but they differ from the more advanced ones.
+Out of the advanced methods it falls out the fastest, and its results are also the best.
+However, it is more difficult to choose second and subsequent places - looking at the results, the simple methods are very quick, but they differ from the more advanced ones.
 Looking at the advanced ones, with the exception of VIM, MissForest stands out.
 However, his problem is time.
-So after VIM, it seems second best to use mean / median for a small amount of time and missForest for a large one.
-However, there was no huge difference in results between the methods - which, however, is not visible in the charts, "removeRows" is characterized by large fluctuations, which we advise against.
+So after VIM, it seems second best to use mean / median replenishment for a small amount of time and missForest for a large one.
+However, although mode/medain usually obtained worse ranks, there was no extensive difference in results between the methods - which is not visible in the charts.
+"removeRows" is characterized by large fluctuations, and therefore we advise against using it.
 
@@
 
-The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a very good package, supplementing the missing values with medians from columns is not at all is such a stupid idea as it may seem and *missForest* is also noteworthy - if we have time for running it and the dataset isn't too huge.
-Unfortunately, it is impossible to draw a wise conclusion based on such a small number of tested data - admittedly as many as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer and assumptions that respectful statistics would praise.
-Nevertheless, above code is certainly very valuable and for more data sets it could confirm one or another belief.
+The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a resplendent package, supplementing the missing values with medians from columns is not is such an unwise idea as it may seem and *missForest* is also noteworthy - if we have time for running it and the dataset isn't too sizeable.
+Unfortunately, it is impossible to draw an objective conclusion based on such a small number of tested data - with as few as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer.
+Nevertheless, the code is certainly very valuable and for more datasets it could confirm one or another belief.
+
 ## References

### 2020-05-30 article.Rmd
spelling

@@
 title: "The Hajada Imputation Test"
-author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
-date: "20/05/2020"
+author: "Jakub Kosterna, Dawid Przybyliński, Hanna Zdulska"
+date: "May 2020"
 output:
@@
 The fact is, there is no one universal best method of imputation, and different problems require various solutions.
-Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us, or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
+Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
 
@@
 
-# Introduction + motivation
+# Introduction and motivation
 
@@
 2. *remove rows* - trivial solution of removing rows containing any missing data.
-3. *'bogo' random replace* - simple algorithm replacing NA values with random numbers or factors from their columns.
+3. *'bogo' random replace* - simple algorithm replacing NA values with random numbers or factors from their features.
 4.
@@
 The fourteen data frames were taken from **OpenML100 collection** and were corrected specifically for this research.
-Both small and simple frames were chosen, as well as more complex and containing a lot of information, difficult to explain or present.
+Both small and simple frames were chosen as well as more complex and containing a lot of information, difficult to explain or present.
 They can be found under the following identifiers with the following names: 1590 - *adult*, 188 - *eucalyptus*, 23381 - *dresses-sales*, 29 - *credit-approval*, 38 - *sick*, 40536 - *SpeedDating*, 41278 - *okcupid-stem*, 56 - *vote*, 6332 - *cylinder-bands*, 1018 - *ipums_la_99-small*, 27 - *colic*, 4 - *labor*, 55 - *hepatitis* and 944 - *echoMonths*.
@@
 
-The six imputations mentioned earlier were conducted on all fourteen of them, but only on six, imputations were successful in their entirety.
+The six imputations mentioned earlier were conducted on all fourteen of them, but only on six imputations were successful in their entirety.
 
@@
 In order to compare the quality of the imputed data and see how this supplementation of deficiencies works in practice, five binary classification algorithms have been selected.
-Their choice was made after careful analysis and extensive discussion, in order to find models that are both widely known and used, and also apply to different approaches and give reasonably distinguishable results using different techniques.
+Their choice was made after careful analysis and extensive discussion, in order to find models that are both widely known and used, but also apply to different approaches and give reasonably distinguishable results using different techniques.
 The final choice fell on:
@@
 Judging by the logarithmic scale, there is no surprise that removing rows and mode/median replenishment are definitely the fastest methods, with removing rows being several times faster.
-what would you expect, also "random values" imputation was quite fast, but slower than the two mentioned.
+What would you expect, also "random values" imputation was quite fast, but slower than the two mentioned.
 Looking at more advanced ones, definitely *VIM* turned up to be the fastest - times usually being approximately 5-10 times shorter than *missForest* and *mice*.
@@
 
-![](files/rankingAll.png)
+![](files/rankingGood.png)
 
@@
 
-![](files/rankingGood.png)
+![](files/rankingAll.png)
 
@@
 So after VIM, it seems second best to use mean / median replenishment for a small amount of time and missForest for a large one.
-However, although mode/medain usually obtained worse ranks, there was no extensive difference in results between the methods - which is not visible in the charts.
+However, although mode / medain usually obtained worse ranks, there was no extensive difference in results between the methods - which is not visible in the charts.
 "removeRows" is characterized by large fluctuations, and therefore we advise against using it.
@@
 
-The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a resplendent package, supplementing the missing values with medians from columns is not is such an unwise idea as it may seem and *missForest* is also noteworthy - if we have time for running it and the dataset isn't too sizeable.
+The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a resplendent package, supplementing the missing values with medians from columns is not is such an unwise idea as it may seem and *missForest* is also noteworthy - if one have time for running it and the dataset isn't too sizeable.
 Unfortunately, it is impossible to draw an objective conclusion based on such a small number of tested data - with as few as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer.

### 2020-05-31 article.Rmd
have -> has

@@
 
-The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a resplendent package, supplementing the missing values with medians from columns is not is such an unwise idea as it may seem and *missForest* is also noteworthy - if one have time for running it and the dataset isn't too sizeable.
+The experiment can be considered successful when it comes to the datasets and tools made available to us - calculated measures, visualizations made, as well as conclusions led us, among others, to the intuition that probably *VIM* is really a resplendent package, supplementing the missing values with medians from columns is not is such an unwise idea as it may seem and *missForest* is also noteworthy - if one has time for running it and the dataset isn't too sizeable.
 Unfortunately, it is impossible to draw an objective conclusion based on such a small number of tested data - with as few as 14 sets, problems with imputations on some of them or too small dataframes not giving a satisfactory answer.

### 2020-06-12 2-2-the-hajada-imputation-test.Rmd
Poprawki do artykułu

--- 2020-05-31 article.Rmd
@@
----
-title: "The Hajada Imputation Test"
-author: "Jakub Kosterna, Dawid Przybyliński, Hanna Zdulska"
-date: "May 2020"
-output:
-  html_document:
-    code_folding: hide
-  pdf_document: default
----
+## The Hajada Imputation Test
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
-getwd()
-```
+*Authors: Jakub Kosterna, Dawid Przybyliński, Hanna Zdulska (Warsaw University of Technology)*
 
-# Abstract
+### Abstract
 
@@
 Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
+In this chapter we will take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
+For this purpose, we will analyse imputation times and machine learning classification models' quality and performance for imputed datasets.
 
-In this chapter which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names) we'll take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
-Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
-For this purpose, we'll test the modified collections for getting rid of missing values on several machine learning classification algorithms, and then considering the selected quality measures after comparing the results, we'll get the desired view.
-Due to the need to obtain a clear and interpretable result, we have limited ourselves to the problem of binary classification.
-
-# Introduction and motivation
+### Introduction and Motivation
 
@@
 Some scientists believe the crux lies in the most advanced and sophisticated algorithms, others trust the simplest of all possible.
+In a way, both sides are right - the first naturally win in speed of execution, but the second ones do not have advanced operations implemented for nothing and this usually results in data that more accurately reflects reality or what might be expected from it.
+How do the individual popular ways fall on the famous collections from OpenML100^1^?
+In this test, which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names), we will find the answer to this question, by delving into how well classification on imputed datasets can perform and when predictions are the best, rank algorithms taken into account and finally choosing the best one.
+Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
 
-## Imputation functions
+#### Imputation functions
 
@@
 
-We'll look at the three simple imputation methods and three more advanced algorithms offered by popular packages.
-The clash of implementations with such a degree of diversity of complexity will give a clear message whether it pays off to reach for an advanced algorithm from the package prepared by professionals for many hours, conducted on many tests, or rather stay with minimalism and write easy code solving the problem in a trivial way.
+Three simple imputation methods and three more advanced algorithms offered by popular packages were taken to consideration and their comparison.
+The clash of implementations with such a degree of diversity of complexity gives a clear message whether it pays off to reach for an advanced algorithm from the package prepared by professionals for many hours, conducted on many tests, or rather stay with minimalism and write easy code solving the problem in a trivial way.
 
@@
 2. *remove rows* - trivial solution of removing rows containing any missing data.
-3. *'bogo' random replace* - simple algorithm replacing NA values with random numbers or factors from their features.
+3. *'bogo' random replace* - simple algorithm replacing each NA value with a random entry from its feature. (chosen independently for every missing value)
 4.
-*mice* ^1^ - advanced method creating multiple imputations for multivariate missing data, based on *Fully Conditional Specification* ^2^, where each incomplete variable is imputed by a separate model.
+*mice* ^2^ - advanced method creating multiple imputations for multivariate missing data, based on *Fully Conditional Specification* ^3^, where each incomplete variable is imputed by a separate model.
 Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets.
-5. *VIM* ^3^ - standard k-nearest neighbors algorithm taken from the library.
-6. *missForest* ^4^ - imputation based on random forest offered by the package.
+5. *VIM* ^4^ - standard k-nearest neighbors algorithm taken from the library.
+6. *missForest* ^5^ - imputation based on random forest offered by the package.
 
-## Datasets
+### Methodology
 
-The fourteen data frames were taken from **OpenML100 collection** and were corrected specifically for this research.
+#### Datasets
+
+The fourteen data frames were taken from OpenML100 collection ^1^ and were corrected specifically for this research.
 Both small and simple frames were chosen as well as more complex and containing a lot of information, difficult to explain or present.
-They can be found under the following identifiers with the following names: 1590 - *adult*, 188 - *eucalyptus*, 23381 - *dresses-sales*, 29 - *credit-approval*, 38 - *sick*, 40536 - *SpeedDating*, 41278 - *okcupid-stem*, 56 - *vote*, 6332 - *cylinder-bands*, 1018 - *ipums_la_99-small*, 27 - *colic*, 4 - *labor*, 55 - *hepatitis* and 944 - *echoMonths*.
+They can be found under the following identifiers with the following names and statistics:
 
-Those above have been placed in individual directories identified by id in the prepared directory.
+|Id     |Name             |   No. rows|  No. columns| No. missings|
+|:------|----------------:|----------:|------------:|------------:|
+|1590   |            adult|      48842|           15|         6465|
+|188    |       eucalyptus|        736|           20|          448|
+|1590   |   dressess-sales|        500|           13|          835|
+|29     |  credit-approval|        690|           16|           67|
+|38     |             sick|       3772|           30|         6064|
+|40536  |      SpeedDating|       8378|          123|        18372|
+|41278  |     okcupid-stem|      50789|           20|       154107|
+|56     |             vote|        435|           17|          392|
+|6332   |   cylinder-bands|        540|           40|          999|
+|1018   |ipums_la_99-small|       8844|           57|        34843|
+|27     |            colic|        368|           23|         1927|
+|4      |            labor|         57|           17|          326|
+|55     |        hepatitis|        155|           20|          167|
+|944    |       echoMonths|        130|           10|           97|
 
-The six imputations mentioned earlier were conducted on all fourteen of them, but only on six imputations were successful in their entirety.
+These above have been placed in individual directories identified by id in the prepared directory.
+The six imputations mentioned earlier were conducted on all fourteen of them, but only for six, all of the methods were successful.
 
-## Binary classification algorithms
+#### Binary classification algorithms
 
@@
 
-* *classification tree from rpart^5^* - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
-* *k-Nearest Neighbors from class^6^ package* - standard knn model attributing a given observation to a target corresponding to its closest observations in space,
-* *naive Bayes from e1071^7^* - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
-* *random forest from ranger^8^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree trying to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
-* *Support Vector Machine from e1071^7^* -  a discriminative classifier formally defined by a separating hyperplane.
+* *classification tree from rpart^6^* - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
+* *k-Nearest Neighbors^7^ from class^8^ package* - standard knn^7^ model attributing a given observation to a target corresponding to its closest observations in space,
+* *naive Bayes from e1071^9^* - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
+* *random forest from ranger^10^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree trying to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
+* *Support Vector Machine from e1071^9^* -  a discriminative classifier formally defined by a separating hyperplane.
 
-# Description of methodology and results
+#### Work automation
 
-## Reading datasets
+In order to automise our code we developed it in such a way so that it is possible to easily conduct further experiments with entirely new datasets, imputation methods and classification models.
+Plenty of time neccesary for imputation was saved using parallel processing^11^.
 
-In order to read all datasets we used another function, which returned a dataframe containing all the important informations about the test matrices.
-The generated frames of imputation results and times were saved using parallel processing^9^.
-
-## Metrics functions
+#### Metrics functions
 
@@
 
-For the final conclusion, ranking concerned only the last one. Additionally, Matthews correlation coefficient measures were also computed.
+For the final conclusion, ranking concerned only the last one. Additionally, Matthews correlation coefficient^12^ measures were also computed.
 
@@
 
-## Comparing imputation times
+### Results
+
+In order to analyse and create results, except for comparing imputations' time and effectiveness for ML models, we need to take into account that not all of imputations were successful for all datasets.
+In further rankings they will have to be considered as the worst for a given dataset.
+
+#### Comparing imputation times
 
@@
 
-![](files/imputationTimesScatter.png)
+For a better view a boxplot was also created:
+
+![](2-2-files/imputationTimesBoxplot.png)
 
@@
 These last two are quite slow, with the former appearing to be slightly faster.
-
-For a better view a boxplot was also created:
-
-![](files/imputationTimesBoxplot.png)
 
@@
 
-Also time was compared for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).
+Also time was compared for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its rows had at least one missing entry).
 Taking into account the mean time, definitely *missForest* was the slowest, but also its standard deviation seemed to be incomparably large - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very sizeable datasets.
@@
 
-## Best measures
+#### Best measures
 
@@
 Then, for each machine learning algorithm a ranking was created - so that the imputation with best results obtained the first rank, second best - 2nd, etc.
-The rank's distributions for every imputation came out as shown by the following boxplot:
+The rank's distributions for every imputation came out as shown by the following boxplot: (rankings omit failed imputations)
 
-![](files/rankingGood.png)
+![](2-2-files/rankingGood.png)
 
@@
 
-![](files/rankingAll.png)
+![](2-2-files/rankingAll.png)
 
-Considering the imputation times and results, it can be said that the knn algorithm from the VIM package proved to be "universally" the best.
+Considering the imputation times and results, it can be said that the knn^7^ algorithm from the VIM package proved to be "universally" the best.
 Out of the advanced methods it falls out the fastest, and its results are also the best.
@@
 
-# Conclusion
+#### Summary and conclusions
 
@@
 
-## References
+#### References
 
-*^1^mice: Multivariate Imputation by Chained Equations, package created by Stef van Buuren, Karin Groothuis-Oudshoorn, Gerko Vink, Rianne Schouten, Alexander Robitzsch, Lisa Doove, Shahab Jolani, Margarita Moreno-Betancur, Ian White, Philipp Gaffert, Florian Meinfelder, Bernie Gray*
+*^1^OpenML100: a predecessor of the OpenML-CC18, a suite of 100 classification datasets from OpenML which were carefully selected to be usable by many algorithms and also represent datasets commonly used in machine learning research*
 
-*^2^Fully conditional specification: Imputes multivariate missing data on a variable-by-variable basis (Van Buuren et al. 2006; Van Buuren 2007a).*
+*^2^mice: Multivariate Imputation by Chained Equations, package created by Stef van Buuren, Karin Groothuis-Oudshoorn, Gerko Vink, Rianne Schouten, Alexander Robitzsch, Lisa Doove, Shahab Jolani, Margarita Moreno-Betancur, Ian White, Philipp Gaffert, Florian Meinfelder, Bernie Gray*
 
-*^3^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
-*^4^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
+*^3^Fully conditional specification: Imputes multivariate missing data on a variable-by-variable basis (Van Buuren et al. 2006; Van Buuren 2007a).*
 
-*^5^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*
+*^4^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
 
-*^6^class: Various functions for classification, including k-nearest neighbour, Learning Vector Quantization and Self-Organizing Maps, package created by Brian Ripley, William Venables*
+*^5^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
 
-*^7^e1071: Misc Functions of the Department of Statistics, Probability Theory Group (Formerly: E1071), TU Wien, package created by David Meyer*
+*^6^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*
 
-*^8^ranger: A Fast Implementation of Random Forests, package created by Marvin N. Wright, Stefan Wager, Philipp Probst*
+*^7^k-nearest neighbors algorithm: a non-parametric method with the input consisting of the k closest training examples in the feature space.*
 
-*^9^parallel: Support for parallel computation, including by forking (taken from package multicore), by sockets (taken from package snow) and random-number generation*
+*^8^class: Various functions for classification, including k-nearest neighbour, Learning Vector Quantization and Self-Organizing Maps, package created by Brian Ripley, William Venables*
 
-*^10^Matthews correlation coefficient: a measure of the quality of binary (two-class) classifications, introduced by biochemist Brian W. Matthews in 1975*
+*^9^e1071: Misc Functions of the Department of Statistics, Probability Theory Group (Formerly: E1071), TU Wien, package created by David Meyer*
+
+*^10^ranger: A Fast Implementation of Random Forests, package created by Marvin N. Wright, Stefan Wager, Philipp Probst*
+
+*^11^parallel: Support for parallel computation, including by forking (taken from package multicore), by sockets (taken from package snow) and random-number generation*
+
+*^12^Matthews correlation coefficient: a measure of the quality of binary (two-class) classifications, introduced by biochemist Brian W. Matthews in 1975*

### 2020-06-18 2-2-the-hajada-imputation-test.Rmd
popraweczki

@@
 In this chapter we will take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
-For this purpose, we will analyse imputation times and machine learning classification models' quality and performance for imputed datasets.
+For this purpose, we will analyse imputation times and predictions scores for several machine learning classification algorithms on various datasets.
 
@@
 In statistics, imputation is the process of replacing missing data with substituted values.
-Over the years, humanity has created many different methods accomplishing that, including the simple and instinctive ones but also more advanced and hard to be easily explained.
+Over the years, humanity has created many different methods accomplishing that, including the simple and instinctive ones, but also more advanced and hard to be easily explained.
 Choosing the best-suited imputation method for the dataset with missing values is still the daily dilemma of every data scientist and there isn't one universally recognized best technique.
@@
 In a way, both sides are right - the first naturally win in speed of execution, but the second ones do not have advanced operations implemented for nothing and this usually results in data that more accurately reflects reality or what might be expected from it.
-How do the individual popular ways fall on the famous collections from OpenML100^1^?
+How do the individual popular ways fall on the famous collections from OpenML100?
 In this test, which we decided to name *The Hajada Imputation Test* (taken from the first two letters of creators' names), we will find the answer to this question, by delving into how well classification on imputed datasets can perform and when predictions are the best, rank algorithms taken into account and finally choosing the best one.
+Hajada Test depends on prediction scores and imputation times.
 Obviously its result will not be the final verdict declaring which imputation algorithm is clearly better or worse, but considering the comparison of their quality on different datasets, we will be able to assess their collective performance and effectiveness against the backdrop of the whole.
+
+### Methodology
 
@@
 4.
-*mice* ^2^ - advanced method creating multiple imputations for multivariate missing data, based on *Fully Conditional Specification* ^3^, where each incomplete variable is imputed by a separate model.
+*mice* [@2-2-mice] - advanced method creating multiple imputations for multivariate missing data, based on *Fully Conditional Specification*, where each incomplete variable is imputed by a separate model.
 Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets.
-5. *VIM* ^4^ - standard k-nearest neighbors algorithm taken from the library.
-6. *missForest* ^5^ - imputation based on random forest offered by the package.
-
-### Methodology
+5. *VIM* [@2-2-VIM] - standard k-nearest neighbors algorithm taken from the library.
+6. *missForest* [@2-2-missForest] - imputation based on random forest offered by the package.
 
@@
 
-The fourteen data frames were taken from OpenML100 collection ^1^ and were corrected specifically for this research.
+The fourteen data frames were taken from OpenML100 collection and were corrected specifically for this research.
 Both small and simple frames were chosen as well as more complex and containing a lot of information, difficult to explain or present.
@@
 
-* *classification tree from rpart^6^* - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
-* *k-Nearest Neighbors^7^ from class^8^ package* - standard knn^7^ model attributing a given observation to a target corresponding to its closest observations in space,
-* *naive Bayes from e1071^9^* - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
-* *random forest from ranger^10^* - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree trying to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
-* *Support Vector Machine from e1071^9^* -  a discriminative classifier formally defined by a separating hyperplane.
+* *classification tree from rpart* [@2-2-rpart] - classic algorithm which uses a decision tree to go from observations about an item to conclusions about the item's target value,
+* *k-Nearest Neighbors from class package* [@2-2-class] - standard knn model attributing a given observation to a target corresponding to its closest observations in space,
+* *naive Bayes from e1071* [@2-2-e1071] - well-known classification computing the conditional a-posterior probabilities of a categorical class variable given independent predictor variables using the Bayes rule.
+* *random forest from ranger* [@2-2-ranger] - algorithm consisting of many decisions trees uses bagging and feature randomness when building each individual tree trying to create an uncorrelated forest of trees whose prediction by committee is more accurate than that of any individual tree.
+* *Support Vector Machine from e1071* [@2-2-e1071] - a discriminative classifier formally defined by a separating hyperplane.
 
@@
 In order to automise our code we developed it in such a way so that it is possible to easily conduct further experiments with entirely new datasets, imputation methods and classification models.
-Plenty of time neccesary for imputation was saved using parallel processing^11^.
+Plenty of time neccesary for imputation was saved using parallel processing.
 
@@
 
-For the final conclusion, ranking concerned only the last one. Additionally, Matthews correlation coefficient^12^ measures were also computed.
+For the final conclusion, ranking concerned only the last one. Additionally, Matthews correlation coefficient measures were also computed.
 
@@
 
-Considering the imputation times and results, it can be said that the knn^7^ algorithm from the VIM package proved to be "universally" the best.
+Considering the imputation times and results, it can be said that the k-nearest neighbors algorithm from the VIM package proved to be "universally" the best.
 Out of the advanced methods it falls out the fastest, and its results are also the best.
@@
 Nevertheless, the code is certainly very valuable and for more datasets it could confirm one or another belief.
-
-#### References
-
-*^1^OpenML100: a predecessor of the OpenML-CC18, a suite of 100 classification datasets from OpenML which were carefully selected to be usable by many algorithms and also represent datasets commonly used in machine learning research*
-
-*^2^mice: Multivariate Imputation by Chained Equations, package created by Stef van Buuren, Karin Groothuis-Oudshoorn, Gerko Vink, Rianne Schouten, Alexander Robitzsch, Lisa Doove, Shahab Jolani, Margarita Moreno-Betancur, Ian White, Philipp Gaffert, Florian Meinfelder, Bernie Gray*
-
-*^3^Fully conditional specification: Imputes multivariate missing data on a variable-by-variable basis (Van Buuren et al. 2006; Van Buuren 2007a).*
-
-*^4^VIM: Visualization and Imputation of Missing Values, package created by Matthias Templ, Alexander Kowarik, Andreas Alfons, Bernd Prantner*
-
-*^5^missForest: Nonparametric Missing Value Imputation using Random Forest Daniel J. Stekhoven*
-
-*^6^rpart: Recursive Partitioning and Regression Trees, package by Terry Therneau, Beth Atkinson, Brian Ripley*
-
-*^7^k-nearest neighbors algorithm: a non-parametric method with the input consisting of the k closest training examples in the feature space.*
-
-*^8^class: Various functions for classification, including k-nearest neighbour, Learning Vector Quantization and Self-Organizing Maps, package created by Brian Ripley, William Venables*
-
-*^9^e1071: Misc Functions of the Department of Statistics, Probability Theory Group (Formerly: E1071), TU Wien, package created by David Meyer*
-
-*^10^ranger: A Fast Implementation of Random Forests, package created by Marvin N. Wright, Stefan Wager, Philipp Probst*
-
-*^11^parallel: Support for parallel computation, including by forking (taken from package multicore), by sockets (taken from package snow) and random-number generation*
-
-*^12^Matthews correlation coefficient: a measure of the quality of binary (two-class) classifications, introduced by biochemist Brian W. Matthews in 1975*

### 2020-06-28 2-2-the-hajada-imputation-test.Rmd
Może nawet ostateczne poprawki

@@
 There are many different ways of dealing with missing values.
-Depending on their quantity and properties, various methods turn out to be the best.
-The fact is, there is no one universal best method of imputation, and different problems require various solutions.
-Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
+Depending on their quantity and properties, various methods turn out to be the best Using one and the same solution for each dataset is certainly not the best option, but sometimes it is convenient or even necessary to focus on one particular method without much insight into its effect - for example, when time limits us or when, due to the complexity of the algorithm or class of problems, we are forced to choose only one method.
 In this chapter we will take a look at several popular imputation methods and try to compare them with each other, including ranking of algorithms taken into account and finally choosing the best one.
